The purpose of this script is to make a map for the manuscript


```{r}
#Clear everything out and start fresh
rm(list=ls())
library(reproducible) #setting up directories for data management
library(dplyr)        #data wranglin'
library(ggplot2)      #data visualization

library(cowplot)
library(googleway)
library(ggrepel)
library(ggspatial)
#library(libwgeom)
library(sf)
library(rnaturalearth) #Map of countries for entire world
library(rnaturalearthdata)
library(geodata)


#This just sets up files for inputs (already created, must contain your files), and outputs
input_directory<-reproducible::checkPath(file.path(getwd(), "0. Map/inputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "0. Map/outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "0. Map/figures"), create = TRUE)


#read in the files we need


#For now just reading in the camera locations from the next folder (sin bin I know)
camloc<-read.csv(file.path(getwd(), "1. Species Response Variables/inputs/OSM_camera_deployment.csv"))

camloc<-camloc%>%
  dplyr::select(array, site, lat, long)


#read in the shape files
lu_polygons<-st_read(file.path(input_directory, "ALL_LUs.shp"))
# If your shapefile isn't in WGS84 (EPSG:4326), we force it to obey our map's projection
lu_polygons <- st_transform(lu_polygons, crs = 4326)

#And remove the ones we dont want
lu_polygons <- lu_polygons %>%
  filter(!(NameRefere %in% c("LU8", "LU4")))
```

Playing with polygons
```{r}
ggplot() +
  # FIRST LAYER: The monitoring areas (Polygons)
  # We use a light fill and a distinct border
  geom_sf(data = lu_polygons, fill = "gray80", color = "black", linewidth = 0.5) +
  
  # SECOND LAYER: Your camera points on top
  geom_sf(data = cam_sf, size = 0.5, color = "red") +
  
  # Keeping your successful top/left axis control
  coord_sf(expand = TRUE, label_axes = list(top = "E", left = "N")) +
  
  theme_bw() +
  theme(
    axis.text.bottom = element_blank(),
    axis.ticks.bottom = element_blank(),
    axis.text.right = element_blank(),
    axis.ticks.right = element_blank(),
    axis.text.top = element_text(size = 9),
    axis.text.left = element_text(size = 9),
    aspect.ratio = 1,
    panel.grid = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)
  ) +
  labs(title = "Monitoring Area Polygons Check", x = NULL, y = NULL)


```




```{r}
# 2. Transform it into a spatial object. 
# Longitude is X, Latitude is Y. Do not mix them up again.
cam_sf <- st_as_sf(camloc, coords = c("long", "lat"), crs = 4326)

# 3. The Basic Map
ggplot(data = cam_sf) +
  geom_sf(size = 0.2, color = "black") +
  coord_sf() + 
  theme_classic()


ggplot(data = cam_sf) +
  # Keep the dots small as you had them
  geom_sf(size = 0.2, color = "black") +
  
  # Set the coordinate system and move the labels
  # 'ndiscr' handles the grid detail; 'default_crs' ensures it knows 4326
  coord_sf(expand = TRUE) +
  
  # Force the box and label positions
  scale_x_continuous(sec.axis = dup_axis(name = NULL)) + # Duplicate X for top line
  scale_y_continuous(sec.axis = dup_axis(name = NULL)) + # Duplicate Y for right line
  
  # Add the scale bar and north arrow in the bottom left (bl)
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"),
                         style = north_arrow_fancy_orienteering) +
  
  # Apply the strict theme
  theme_bw() + 
  theme(
    # Remove the labels from the bottom and right
    axis.text.bottom = element_blank(),
    axis.ticks.bottom = element_blank(),
    axis.text.right = element_blank(),
    axis.ticks.right = element_blank(),
    # Ensure top and left are visible
    axis.text.top = element_text(size = 9),
    axis.text.left = element_text(size = 9),
    # Make it a perfect square box
    aspect.ratio = 1,
    panel.grid.major = element_blank(), # Clean background
    panel.grid.minor = element_blank()
  ) +
  labs(x = NULL, y = NULL) # Clean up the titles


main_plot<-ggplot(data = cam_sf) +
  # Small points, just as you requested
  geom_sf(size = 0.2, color = "black") +
  
  # DIRECT COMMAND to coord_sf to move the labels
  # N = North (Latitude), E = East (Longitude)
  coord_sf(expand = TRUE, label_axes = list(top = "E", left = "N")) +
  
  # Scale bar and North arrow
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"),
                         style = north_arrow_fancy_orienteering) +
  
  # The strict square theme
  theme_bw() + 
  theme(
    # Only the Top and Left labels are permitted to display
    # We turn off the others here
    axis.text.bottom = element_blank(),
    axis.ticks.bottom = element_blank(),
    axis.text.right = element_blank(),
    axis.ticks.right = element_blank(),
    
    # Ensure the top and left remain legible
    axis.text.top = element_text(size = 9),
    axis.text.left = element_text(size = 9),
    
    # Square geometry and clean background
    aspect.ratio = 1,
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  labs(x = NULL, y = NULL)

main_plot<-ggplot(data = cam_sf) +
  
  geom_sf(data = lu_polygons, fill = "darkgreen", color = "black", linewidth = 0.5) +
  # Small points, exactly as you wanted them
  geom_sf(size = 0.2, color = "black") +
  
  # Keeping your successful coordinate commands
  coord_sf(expand = TRUE, label_axes = list(top = "E", left = "N")) +
  
  # Scale bar and North arrow
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"),
                         style = north_arrow_fancy_orienteering) +
  
  # The strict square theme
  theme_bw() + 
  theme(
    # Command the background to be grey
    panel.background = element_rect(fill = "gray90"),
    
    # Ensure the border is a sharp black box
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    
    # Silence the redundant labels
    axis.text.bottom = element_blank(),
    axis.ticks.bottom = element_blank(),
    axis.text.right = element_blank(),
    axis.ticks.right = element_blank(),
    
    # Keep the top and left legible
    axis.text.top = element_text(size = 9),
    axis.text.left = element_text(size = 9),
    
    # Geometry perfection
    aspect.ratio = 1,
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  labs(x = NULL, y = NULL)


ggplot(data = cam_sf) +
  # 1. Map the aesthetics with a shared 'Legend' name in the scales
  geom_sf(data = lu_polygons, aes(fill = "Monitoring Areas"), color = "black", linewidth = 0.5) +
  geom_sf(aes(color = "Camera Sites"), size = 0.2) +
  
  # 2. FORCE them together by using identical 'name' and 'guide' arguments
  scale_fill_manual(values = c("Monitoring Areas" = "darkgreen"), name = "Legend") +
  scale_color_manual(values = c("Camera Sites" = "black")) +
  
  # 3. Explicitly tell ggplot to merge them into one guide
  guides(
    fill = guide_legend(order = 1),
    color = guide_legend(order = 2)
  ) +
  
  coord_sf(expand = TRUE, label_axes = list(top = "E", left = "N")) +
  
  # Scale and North Arrow
  annotation_scale(location = "bl", width_hint = 0.3) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(1.75, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering) +
  
  theme_bw() + 
  theme(
    panel.background = element_rect(fill = "gray90"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    
    # Legend aesthetics
    legend.position = c(0.02, 0.065),
    legend.justification = c(0, 0),
    legend.box.background = element_rect(fill = "white", color = "black"),
    legend.key = element_blank(),
    legend.spacing.y = unit(0, "pt"), # Remove vertical gap between items
    legend.margin = margin(2, 5, 2, 2), # Tighten internal padding
    
    axis.text.bottom = element_blank(),
    axis.ticks.bottom = element_blank(),
    axis.text.right = element_blank(),
    axis.ticks.right = element_blank(),
    axis.text.top = element_text(size = 9),
    axis.text.left = element_text(size = 9),
    aspect.ratio = 1,
    panel.grid = element_blank()
  ) +
  labs(x = NULL, y = NULL)

ggplot(data = cam_sf) +
  # 1. Map aesthetics
  geom_sf(data = lu_polygons, aes(fill = "Monitoring Areas"), color = "black", linewidth = 0.5) +
  geom_sf(aes(color = "Camera Sites"), size = 0.2) +
  
  # 2. MATCH THE NAMES EXACTLY to merge the boxes under one "Legend" title
  scale_fill_manual(values = c("Monitoring Areas" = "darkgreen"), name = "Legend") +
  scale_color_manual(values = c("Camera Sites" = "black"), name = "Legend") +
  
  # 3. Explicitly merge the guides
  guides(
    fill = guide_legend(order = 1),
    color = guide_legend(order = 2)
  ) +
  
  coord_sf(expand = TRUE, label_axes = list(top = "E", left = "N")) +
  
  # Scale and North Arrow
  annotation_scale(location = "bl", width_hint = 0.3) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(1.75, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering) +
  
  theme_bw() + 
  theme(
    panel.background = element_rect(fill = "gray90"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    
    # Legend aesthetics - sitting perfectly in the corner
    legend.position = c(0.02, 0.065),
    legend.justification = c(0, 0),
    legend.background = element_blank(), # Let the box background handle it
    legend.box.background = element_rect(fill = "white", color = "black"),
    legend.key = element_blank(),
    legend.spacing.y = unit(0, "pt"), 
    legend.margin = margin(2, 5, 2, 2), 
    
    axis.text.bottom = element_blank(),
    axis.ticks.bottom = element_blank(),
    axis.text.right = element_blank(),
    axis.ticks.right = element_blank(),
    axis.text.top = element_text(size = 9),
    axis.text.left = element_text(size = 9),
    aspect.ratio = 1,
    panel.grid = element_blank()
  ) +
  labs(x = NULL, y = NULL)


main_plot<-ggplot() +
  # Polygon layer with manual fill
  geom_sf(data = lu_polygons, aes(fill = "Monitoring Areas"), color = "black", linewidth = 0.5) +
  
  # Point layer with manual color (we'll treat it as "fill" for legend merging)
  geom_sf(data = cam_sf, aes(fill = "Camera Sites"), color = "black", size = 0.2) +
  
  # Single legend with both items
  scale_fill_manual(
    name = "Legend",
    values = c(
      "Monitoring Areas" = "darkgreen",
      "Camera Sites" = "black"
    )
  ) +
  
  guides(
    fill = guide_legend(
      override.aes = list(
        color = c("black", "black"),  # ensures points have black border
        shape = c(22, 21),            # different shapes for polygon vs point
        size = c(5, 2)                # size for points vs polygons
      )
    )
  ) +
  
  coord_sf(expand = TRUE, label_axes = list(top = "E", left = "N")) +
  annotation_scale(location = "bl", width_hint = 0.3) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(1.75, "in"), pad_y = unit(0.5, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "gray90"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    legend.position = c(0.02, 0.065),
    legend.justification = c(0, 0),
    legend.background = element_blank(),
    legend.box.background = element_rect(fill = "white", color = "black"),
    legend.key = element_blank(),
    legend.spacing.y = unit(0, "pt"),
    legend.margin = margin(2, 5, 2, 2),
    axis.text.bottom = element_blank(),
    axis.ticks.bottom = element_blank(),
    axis.text.right = element_blank(),
    axis.ticks.right = element_blank(),
    axis.text.top = element_text(size = 9),
    axis.text.left = element_text(size = 9),
    aspect.ratio = 1,
    panel.grid = element_blank()
  ) +
  labs(x = NULL, y = NULL)
#Next challenge
# 1. Calculate the bounding box (extent) of your camera locations
# We'll turn it into a polygon so we can draw it on the Alberta map
camera_extent <- st_as_sfc(st_bbox(cam_sf))

# 2. Create the Inset Map: Alberta with a red extent box
inset_map <- ggplot() +
  geom_sf(data = st_as_sf(alberta_bg), fill = "gray40", color = "gray2") +
  geom_sf(data = camera_extent, fill = NA, color = "red", linewidth = 1) +
  theme_void()+
  theme(
    # This creates the box around the inset itself
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)
  )# Removes all axes and labels for a clean inset


inset_map <- ggplot() +
  geom_sf(data = st_as_sf(alberta_bg), fill = "gray50", color = "black") +
  geom_sf(data = camera_extent, fill = NA, color = "red", linewidth = 1) +
  theme_void() +
  theme(
    plot.background = element_blank(),
    plot.margin = margin(0, 0, 0, 0, "pt"),
    panel.border = element_blank() # Remove the internal border
  )

# 4. Use cowplot to put the inset on top of the main map
# Adjust x, y, width, and height to place it exactly where I want it
ggdraw() +
  draw_plot(main_plot) +
  draw_plot(inset_map, x = 0.65, y = 0.65, width = 0.25, height = 0.25)

map<-ggdraw() +
  draw_plot(main_plot) +
  # This adds a crisp, uniform frame without the 'thick line' glitch
  geom_rect(aes(xmin = 0.72, xmax = 0.84, ymin = 0.44, ymax = 0.75), 
            color = "black", fill = "white", linewidth = 0.6)+
  # We place the inset and then draw a clean box over its coordinates
  draw_plot(inset_map, x = 0.68, y = 0.45, width = 0.2, height = 0.28) +
  draw_label("", x = 0.73, y = 0.7, hjust = 0, vjust = 0)  # Anchor for the box
  

ggsave("Final_Monitoring_Map.png", 
       plot = map, 
       width = 8, 
       height = 8, 
       units = "in", 
       dpi = 300)


```


Mapping tutorial since I am new to this!
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

ggplot(data = world) +
    geom_sf()

ggplot(data = world) +
    geom_sf() +
    annotation_scale(location = "bl", width_hint = 0.5)+
  annotation_north_arrow(location = "bl", which_north = "true", 
       #Change position here
        pad_x = unit(0.90, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering)


install.packages("devtools")
devtools::install_github("ropensci/rnaturalearthhires")



# Download level 1 (provinces) for Canada
canada_sf <- gadm(country="CAN", level=1, path=tempdir())
alberta_bg <- canada_sf[canada_sf$NAME_1 == "Alberta", ]

# Plotting is the same
ggplot(st_as_sf(alberta)) +
  geom_sf() +
  coord_sf()+
    annotation_scale(location = "bl", width_hint = 0.5)+
  annotation_north_arrow(location = "bl", which_north = "true", 
       #Change position here
        pad_x = unit(0.60, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering)+
  theme_classic()

```




```{r}
#Quickly widdle down the camloc file
camloc<-camloc%>%
  dplyr::select(array, site, lat, long)

```

```{r}
# ==============================
# Camera locations map (CartoDB)
# ==============================

# Packages
library(sf)
library(ggplot2)
library(dplyr)
library(ggspatial)
library(geodata)

# ------------------------------
# 1. Convert camera table to sf
# ------------------------------
cams_sf <- st_as_sf(
  camloc,
  coords = c("long", "lat"),
  crs = 4326   # WGS84 lat/long
)

# ------------------------------
# 2. Choose CRSs
# ------------------------------
analysis_crs <- 3400  # NAD83 / Alberta 10-TM Forest
plot_crs     <- 3857  # Web Mercator (required for map tiles)

cams_sf <- st_transform(cams_sf, analysis_crs)

# ------------------------------
# 3. Pull Alberta boundary (GADM)
# ------------------------------
can_prov <- gadm(
  country = "CAN",
  level = 1,
  path = tempdir()
)

alberta <- st_as_sf(can_prov) %>%
  st_transform(analysis_crs) %>%
  filter(NAME_1 == "Alberta")

# ------------------------------
# 4. Transform to plotting CRS
# ------------------------------
cams_plot    <- st_transform(cams_sf, plot_crs)
alberta_plot <- st_transform(alberta, plot_crs)

# ------------------------------
# 5. Define map extent
# ------------------------------
bbox <- st_bbox(cams_plot)

bbox[c("xmin", "ymin")] <- bbox[c("xmin", "ymin")] - 50000
bbox[c("xmax", "ymax")] <- bbox[c("xmax", "ymax")] + 50000

# ------------------------------
# 6. Plot
# ------------------------------
ggplot() +
  
  # Grey basemap tiles
  annotation_map_tile(
    type = "cartodb_positron",
    zoom = 8
  ) +
  
  # Alberta outline (subtle)
  geom_sf(
    data = alberta_plot,
    fill = NA,
    color = "grey60",
    linewidth = 0.4
  ) +
  
  # Camera locations
  geom_sf(
    data = cams_plot,
    color = "grey20",
    size = 1.6,
    alpha = 0.9
  ) +
  
  # Zoom to study area
  coord_sf(
    xlim = c(bbox["xmin"], bbox["xmax"]),
    ylim = c(bbox["ymin"], bbox["ymax"]),
    expand = FALSE
  ) +
  
  # Scale bar
  annotation_scale(
    location = "bl",
    width_hint = 0.25
  ) +
  
  # North arrow ABOVE scale bar
  annotation_north_arrow(
    location = "bl",
    which_north = "true",
    pad_y = unit(1.3, "cm"),
    pad_x = unit(0.5, "cm"),
    style = north_arrow_fancy_orienteering
  ) +
  
  # Clean theme
  theme_void()




```



```{r}

#Convert the points to an sf object
cams_sf <- st_as_sf(
  camloc,
  coords = c("long", "lat"),
  crs = 4326   # WGS84 (lat/long)
)

#Choose a projected CRS for northern Alberta
map_crs <- 3400  # NAD83 / Alberta 10-TM Forest
cams_sf <- st_transform(cams_sf, map_crs)

#Pull a basemap
canada <- ne_countries(
  country = "canada",
  scale = "medium",
  returnclass = "sf"
) %>%
  st_transform(map_crs)

#Pull Alberta
# Download Canadian provinces (level 1)
can_prov <- gadm(
  country = "CAN",
  level = 1,
  path = tempdir()
)

can_prov <- st_as_sf(can_prov) %>%
  st_transform(map_crs)

# Extract Alberta
alberta <- can_prov %>% filter(NAME_1 == "Alberta")


bbox <- st_bbox(cams_sf)

bbox_expanded <- bbox
bbox_expanded[c("xmin", "ymin")] <- bbox[c("xmin", "ymin")] - 50000
bbox_expanded[c("xmax", "ymax")] <- bbox[c("xmax", "ymax")] + 50000
```


```{r}
ggplot() +
  
  # Alberta background
  geom_sf(
    data = alberta,
    fill = "grey95",
    color = "grey70",
    linewidth = 0.3
  ) +
  
  # Camera locations
  geom_sf(
    data = cams_sf,
    color = "grey30",
    size = 1.5,
    alpha = 0.8
  ) +
  
  # Zoom to study area
  coord_sf(
    xlim = c(bbox_expanded["xmin"], bbox_expanded["xmax"]),
    ylim = c(bbox_expanded["ymin"], bbox_expanded["ymax"]),
    expand = FALSE
  ) +
  
  # Scale bar + north arrow
  annotation_scale(
    location = "bl",
    width_hint = 0.3,
    text_cex = 0.9
  ) +
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  
  # Clean theme
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_text(size = 10)
  )

```


Lets prepare the data for modeling.
```{r}

#PC variables are already scaled, but lets scale the remaining variables
wtd_dets<-wtd_dets%>%
  mutate(across(c(disturbance_index, polygonal_index, linear_index, mean_snow_spatial, snow_anomaly, predator_index, lat), .fns = ~ as.numeric(scale(.x))))

         
#Lets do a quick check of correlations
vars <- wtd_dets %>%
  dplyr::select(disturbance_index, polygonal_index, linear_index, mean_snow_spatial, snow_anomaly, predator_index, lat, PC1, PC2)

# Compute correlation matrix (use complete.obs to ignore NAs)
cor_matrix <- cor(vars, use = "complete.obs")


#And now lets identify the actual correlation values to report in the manuscrpt
cor_df <- as.data.frame(cor_matrix) %>%
  tibble::rownames_to_column(var = "var1") %>%
  pivot_longer(
    cols = -var1,
    names_to = "var2",
    values_to = "correlation"
  )%>%
  # Remove self-correlations and duplicate pairs easy peasy (or is it peazy?)
  filter(var1 < var2)%>%
   # Keep only strong correlations
  filter(abs(correlation) >= 0.5)%>%
  arrange(desc(abs(correlation)))

#Yeah so there is a strong correlation between polygonal features and total disturbance, this makes sense right, since if you think about it, the disturbance measure is the sum of all polygonal and linear features. And since linear features are such a small footprint, the majority of the total disturbance is just polygonal. So we wont model total disturbance, and make sure we talk about that in the manuscript

#What is the actual value to report though
cor.test(wtd_dets$disturbance_index, wtd_dets$polygonal_index)


# Print the correlation matrix now without disturbance index
vars <- wtd_dets %>%
  dplyr::select(#disturbance_index, 
                polygonal_index, linear_index, mean_snow_spatial, snow_anomaly, predator_index, lat, PC1, PC2)
cor_matrix <- cor(vars, use = "complete.obs")
print(round(cor_matrix, 2))

#Produce the plot and save it
png(file.path(figure_directory, "correlation_plot.png"))
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, addCoef.col = "black")  
dev.off()


#Write that dataframe with the scaled variables to the outputs folder (this will be used in the next step for modeling)
write.csv(wtd_dets, file.path(output_directory, "final_df.csv"), row.names = FALSE)

```








Now for males
```{r}

#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_males)

#What about a "null" model that doesn't include disturbance?
null_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_males)



#All disturbance
# disturbance_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
#                    disturbance_index+
#                    PC1 + PC2 + #Habitat covariates
#                    predator_index + #Predator index
#                    lat+
#                    mean_snow_spatial*disturbance_index+
#                    snow_anomaly*disturbance_index+
#                    (1|array),
#                  family = "binomial",
#                  data = wtd_dets)
# summary(disturbance_males)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_males)


#Linear
linear_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_males)



#Model selection and cleanup into a nice dataframe!
model_selection_male<- as.data.frame(model.sel(global_males, null_males, linear_males, polygonal_males,
              rank = "AICc"))

model_selection_male<-model_selection_male%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_male)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_male)<-NULL




```


Females
```{r}
#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_females)

#What about a "null" model that doesn't include disturbance?
null_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_females)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_females)


#Linear
linear_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_females)



#Model selection and cleanup into a nice dataframe!
model_selection_females<- as.data.frame(model.sel(global_females, null_females, linear_females, polygonal_females,
              rank = "AICc"))

model_selection_females<-model_selection_females%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_females)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_females)<-NULL




```


Young Deer?
```{r}
#All disturbance
#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_young)

#What about a "null" model that doesn't include disturbance?
null_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_young)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_young)


#Linear
linear_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_young)



#Model selection and cleanup into a nice dataframe!
model_selection_young<- as.data.frame(model.sel(global_young, null_young, linear_young, polygonal_young,
              rank = "AICc"))

model_selection_young<-model_selection_young%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_young)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_young)<-NULL



```

Nicely format those model seleciton tables, this is a key result!
```{r}
model_selection_male
model_selection_females
model_selection_young


#--- Helper function to clean each table ---
clean_table <- function(df, group_name) {
  df %>%
    mutate(
      group = group_name,

      # Extract just the model part before the underscore
      model = str_extract(model, "^[^_]+"),

      # Convert to nice names
      model = factor(model,
                     levels = c("linear", "polygonal", "global", "null"),
                     labels = c("Linear", "Polygonal", "Global", "Null"))
    )
}

# Clean all three
male_clean    <- clean_table(model_selection_male,    "Males")
female_clean  <- clean_table(model_selection_females, "Lone Females")
young_clean   <- clean_table(model_selection_young,   "Females With Young")

combined_models <- bind_rows(male_clean, female_clean, young_clean)%>%
  relocate(group)%>%
  mutate(logLik = round(logLik, digits = 2),
         AICc = round(AICc, digits = 2),
         delta = round(delta, digits = 2))




library(flextable)
library(officer)

# Convert to a nice Word table
ft <- flextable::flextable(combined_models)

# Optimize the widths and heights of columns. Nice to have!
ft <- flextable::autofit(ft)

#And then turn that into an actual word document
doc <- officer::read_docx()
doc <- flextable::body_add_flextable(doc, ft)

#Now actually write the file
print(doc, target = file.path(output_directory, "model_selection_results.docx"))

```




Ok, in the following here is what I want to do. I want to write a loop that performs an out of sample test of our model selection, to identify where the "best" top model is identified
```{r}

#These are the response variables to iterate over
response_variables<-data.frame(response = c("cbind(n_male, (n_weeks - n_male)) ~",
                      "cbind(n_female, (n_weeks - n_female)) ~",
                      "cbind(n_young, (n_weeks - n_young)) ~"),
                      name = c("males", "females", "young"))


#Create a blank list of model selection tables
model_selection_tables<-list()

#How many simulations to do?
n_sims<-100


#Random seed
set.seed(420)
for(j in 1:nrow(response_variables)){
    
  #Store the current response variable we are working on 
  response <- as.character(response_variables[j, 1])
  #And the name, we will need that for later
  name<- as.character(response_variables[j, 2])
  
  print(paste("Working on models for", name, "at", Sys.time()))
  
  
  #Pre-emptive, but its more efficient to define the model formulas here within this loop, so its only done once per group rather than everytime within the next loop! #Efficiency 101 baby
  formula_global<- as.formula(paste(response,
                      "polygonal_index+linear_index+PC1 + PC2 +predator_index +lat+mean_snow_spatial+snow_anomaly+
                           + (1|array)", sep = ""))
  
  formula_null<- as.formula(paste(response,
                      "PC1 + PC2 +predator_index +lat+mean_snow_spatial+snow_anomaly+
                           + (1|array)", sep = "")) 
  formula_linear<- as.formula(paste(response,
                      "linear_index+
                      PC1 + PC2 +predator_index +lat+
                      mean_snow_spatial*linear_index+
                      snow_anomaly*linear_index+
                           + (1|array)", sep = ""))
  formula_polygonal<- as.formula(paste(response,
                      "polygonal_index+
                      PC1 + PC2 +predator_index +lat+
                      mean_snow_spatial*polygonal_index+
                      snow_anomaly*polygonal_index+
                           + (1|array)", sep = ""))
  
  
  #Now that we have identified our response variable, we need to many
  #many times sample the data, fit each model, run model selection,
  #and identify the top model. 
  for(i in 1:n_sims){
    
    #Sample the dataframe - THIS IS WHERE WE DEFINE THE SAMPLING AMOUNT
    sample<-wtd_dets%>%
      slice_sample(prop = 0.90)
    
    
    #Fit each model!
    fit_global<-glmmTMB::glmmTMB(
                formula = formula_global,
                 family = "binomial",
                 data = sample)
    
    fit_null<-glmmTMB::glmmTMB(
                formula = formula_null,
                 family = "binomial",
                 data = sample)
    
    fit_linear<-glmmTMB::glmmTMB(
                formula = formula_linear,
                 family = "binomial",
                 data = sample)
    
    fit_polygonal<-glmmTMB::glmmTMB(
                formula = formula_polygonal,
                 family = "binomial",
                 data = sample)
    
    #Quick model selection! and cleanup
   model_sel<-as.data.frame(MuMIn::model.sel(fit_global, fit_null, fit_linear, fit_polygonal,
              rank = "AICc"))
   
   model_sel<-model_sel%>%
    # Add model names as a column
  mutate(model = rownames(model_sel)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))
   
   #Get rid of the rownames, they are annoying
   rownames(model_sel)<-NULL
   
   #Add the name, and the simulation number to this dataframe
   model_sel<-model_sel%>%
     mutate(group = name,
            simulation = i)%>%
     relocate(group, simulation)
   
   # Initialize the nested list if this is the first simulation for this group
if (! name %in% names(model_selection_tables)) {
  model_selection_tables[[name]] <- list()
}

# Store the model table for this simulation
model_selection_tables[[name]][[i]] <- model_sel
  }
  
  #Nothing to see here!
  
}
```


And here we are formatting the final figure and whatnot
```{r}
#Ok now lets take those results and strip them out into a single dataframe
all_model_results <- purrr::map_df(
  model_selection_tables,
  ~ purrr::map_df(.x, ~ .x)
)

#And finally, for each group, how often was the top model top?
all_models <- c("fit_linear", "fit_polygonal", "fit_global", "fit_null")


summary_table <-
  all_model_results %>%
  # Only the top model per simulation
  filter(delta == 0) %>%
  group_by(group, model) %>%
  summarise(n_obs = n(), .groups = "drop") %>%
  
  # Expand so every group has all 4 models
  tidyr::complete(
    group,
    model = all_models,
    fill = list(n_obs = 0)
  ) %>%
  
  # Add total sims and proportion
  mutate(
    total = n_sims,
    prop_top = n_obs / total
  )

#And a plot!
# Set factor order
summary_table <- summary_table %>%
  mutate(
    group = factor(group,
                   levels = c("males", "females", "young"),
                   labels = c("Males", "Lone Females", "Females With Young")),
    model = factor(model,
                   levels = c("fit_linear", "fit_polygonal", "fit_global", "fit_null"),
                   labels = c("Linear", "Polygonal", "Global", "Null")))

#Define the colour blind paletee 
okabe_ito <- c(
  "Linear"     = "#0072B2",
  "Polygonal"  = "#E69F00",
  "Global"     = "#009E73",
  "Null"       = "#D55E00"
)

#Save that figure
model_selection_validation_plot<-ggplot(summary_table, aes(x = model, y = prop_top, fill = model)) +
  geom_col(width = 0.7, alpha = 0.9) +
  
  # Add bar labels
  geom_text(aes(label = round(prop_top, 2)),
            vjust = -0.3,            # position above bar
            size = 4,                # text size
            fontface = "bold") +     # optional
  
  facet_wrap(~ group) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     limits = c(0, 1)) +
  scale_fill_manual(values = okabe_ito) +
  labs(
    title = "Proportion of Samples in Which Each Model Was Top",
    x = "Model",
    y = "Proportion Top Model"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )


#Save that figure
#Save that figure 
ggsave(filename = "model_selection_validation_plot.jpeg", #name
        plot = model_selection_validation_plot,  #plot you want to save
        path = figure_directory, #where its saved
        width = 3000, #how wide
        height = 2000,#how tall
         units = "px", #units in pixels
         bg = 'white') #make sure background is white
```
