The purpose of this script is to make a quick map for the manuscript. This uses the camera locations from the next step, as well as shape files for the landscape units. Nothing fancy to see here.

January 21/25
Andrew Barnas
andrewbarnas@uvic.ca

Load the packages and data we will need
```{r}
#Clear everything out and start fresh
rm(list=ls())
library(reproducible) #setting up directories for data management
library(dplyr)        #data wranglin'
library(ggplot2)      #data visualization
library(cowplot)      #Arranging multiple map elements
library(ggspatial)    #spatial stuff
library(sf)           #real spatial stuff
#library(rnaturalearth) #Map of countries for entire world
#library(rnaturalearthdata)
library(geodata)


#This just sets up files for inputs (already created, must contain your files), and outputs
input_directory<-reproducible::checkPath(file.path(getwd(), "0. Map/inputs"), create = TRUE)
#Dont need an output directory, just a figure directory!
figure_directory<-reproducible::checkPath(file.path(getwd(), "0. Map/figures"), create = TRUE)


#read in the files we need
#The camera locations from the next step (which was the original first step, sue me)
camloc<-read.csv(file.path(getwd(), "1. Species Response Variables/inputs/OSM_camera_deployment.csv"))

#Edit that file down for simplicity
camloc<-camloc%>%
  dplyr::select(array, site, lat, long)

#Transform camera locations into a spatial object. 
cam_sf <- st_as_sf(camloc, coords = c("long", "lat"), crs = 4326)

#read in the shape files
lu_polygons<-st_read(file.path(input_directory, "ALL_LUs.shp"))
# If your shapefile isn't in WGS84 (EPSG:4326), set it as such
lu_polygons <- st_transform(lu_polygons, crs = 4326)

#And remove the LUs we dont want! 
lu_polygons <- lu_polygons %>%
  filter(!(NameRefere %in% c("LU8", "LU4")))
```


Create the actual map, this involves a bit of processing of the data, and then carefully layering elements onto it before saving
```{r}



main_plot<-ggplot() +
  # Polygon layer with manual fill
  geom_sf(data = lu_polygons, aes(fill = "Monitoring Areas"), color = "black", linewidth = 0.5) +
  
  # Point layer with manual color (we'll treat it as "fill" for legend merging)
  geom_sf(data = cam_sf, aes(fill = "Camera Sites"), color = "gray18", shape =16, size = 0.2) +
  
  # Single legend with both items
  scale_fill_manual(
    name = "Legend",
    values = c(
      "Monitoring Areas" = "darkseagreen",
      "Camera Sites" = "gray18"
    )
  ) +
  
  guides(
    fill = guide_legend(
      override.aes = list(
        color = c("black", "black"),  # ensures points have black border
        shape = c(16, 22),            # different shapes for polygon vs point
        size = c(5, 3)                # size for points vs polygons
      )
    )
  ) +
  
  coord_sf(expand = TRUE, label_axes = list(top = "E", left = "N")) +
  annotation_scale(location = "bl", width_hint = 0.3) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(1.5, "in"), pad_y = unit(0.4, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "gray90"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    legend.position = c(0.02, 0.065),
    legend.justification = c(0, 0),
    legend.background = element_blank(),
    legend.box.background = element_rect(fill = "white", color = "black"),
    legend.key = element_blank(),
    legend.spacing.y = unit(0, "pt"),
    legend.margin = margin(2, 5, 2, 2),
    axis.text.bottom = element_blank(),
    axis.ticks.bottom = element_blank(),
    axis.text.right = element_blank(),
    axis.ticks.right = element_blank(),
    axis.text.top = element_text(size = 9),
    axis.text.left = element_text(size = 9),
    aspect.ratio = 1,
    panel.grid = element_blank()
  ) +
  labs(x = NULL, y = NULL)
#Next challenge
# 1. Calculate the bounding box (extent) of your camera locations
# We'll turn it into a polygon so we can draw it on the Alberta map
camera_extent <- st_as_sfc(st_bbox(cam_sf))


#Pull the alberta province layer
canada_provs <- gadm(country = "CAN", level = 1, path = ".")

# 2. Convert to 'sf' format and filter for Alberta
# Note: GADM typically uses 'NAME_1' for the province name
alberta_bg <- st_as_sf(canada_provs) %>%
  filter(NAME_1 == "Alberta")



# 2. Create the Inset Map: Alberta with a red extent box
inset_map <- ggplot() +
  geom_sf(data = st_as_sf(alberta_bg), fill = "gray50", color = "black") +
  geom_sf(data = camera_extent, fill = NA, color = "red", linewidth = 1) +
  theme_void() +
  theme(
    plot.background = element_blank(),
    plot.margin = margin(0, 0, 0, 0, "pt"),
    panel.border = element_blank() # Remove the internal border
  )

# 4. Use cowplot to put the inset on top of the main map
# Adjust x, y, width, and height to place it exactly where I want it
map<-ggdraw() +
  draw_plot(main_plot) +
  # This adds a crisp, uniform frame without the 'thick line' glitch
  geom_rect(aes(xmin = 0.72, xmax = 0.84, ymin = 0.44, ymax = 0.75), 
            color = "black", fill = "white", linewidth = 0.6)+
  # We place the inset and then draw a clean box over its coordinates
  draw_plot(inset_map, x = 0.68, y = 0.45, width = 0.2, height = 0.28) +
  draw_label("", x = 0.73, y = 0.7, hjust = 0, vjust = 0)  # Anchor for the box
  
#Exporting this figure seems to be a nightmare, it moves all the elements around....so out of laziness I am just going to copy and paste. If any students see this, do as I say, not as I do.


```
