The purpose of this file is to do the setup of the response varible for deer? Lets do this clean. 


```{r}
#Clear everything out and start fresh
rm(list=ls())
library(reproducible) #setting up directories for data management
library(readr)        #Needed to read and merge all the csv's in the camera folder
library(dplyr)        #data wranglin'
library(lubridate)    #datetime manipulation
library(tidyr)        #more data manipulation
library(stringr)      #detecting character strings
library(ggplot2)      #data visualization
library(purrr)        #feline-friendly loops 
library(corrplot)     #correlation tests


#This just sets up files for inputs (already created, must contain your files), and outputs
input_directory<-reproducible::checkPath(file.path(getwd(), "inputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "figures"), create = TRUE)


#read in the files we need
dets<-read.csv(file.path(input_directory, "OSM_camera_detections.csv"))
camop<-read.csv(file.path(input_directory, "OSM_camera_deployment.csv"))
covs<-read.csv(file.path(input_directory, "OSM_camera_covariates.csv"))

#And the weather data
weather<-read.csv(file.path(input_directory, "daymet_weather_data.csv"))
weather_10<-read.csv(file.path(input_directory, "daymet_weather_data_10year.csv"))


```


First, lets explore when and where our detections of deer were. 
```{r}
#First, I want to know when deer were detected, this is going to influence when and where we look for deer
wtd<-dets%>%
  filter(species == "White-tailed deer")%>%
  #Just ensuring its read in as a datetime here!
  mutate(datetime = ymd_hms(datetime))


#arrange the detection by site, species, and time, then group by site
wtd<-wtd%>%
  arrange(array, site, datetime)%>%
  group_by(site)%>%
  #Create a lag time that will be used to calculate time differences
   mutate(duration = as.numeric(difftime(datetime,lag(datetime),units = "mins")))

#Set up the framework for assigning unique IDs
wtd$Event.ID<-9999 #This is just an arbitrary event identification code
mins<-2 # THIS IS THE DETECTION EVENT BREAK-POINT YOU CAN CHANGE
seq <- as.numeric(paste0(nrow(wtd),0))
seq <- round(seq,-(nchar(seq)))

#This will go through each row and figure out how far apart in time each one is
#So this creates unique event identifications for subsequent rows IF they are more than 30mins apart
for (i in 2:nrow(wtd)) {
 wtd$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F))
  if(is.na(wtd$duration[i]) | abs(wtd$duration[i]) > mins){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(wtd$duration[nrow(wtd)] < mins| 
   is.na(wtd$duration[nrow(wtd)])){
  wtd$Event.ID[nrow(wtd)] <- wtd$Event.ID[nrow(wtd)-1] 
} else{
  wtd$Event.ID[nrow(wtd)] <- paste0("E",format(seq+1, scientific = F)) 
}

#Lets reduce this down to the information we want (too many columns to bring along!)
wtd<-wtd%>%
  dplyr::select(array, site, datetime, female, yoy, Event.ID)

#Second, since we don't care about absolute number of deer in events, we are going to turn things into binary present or absent. 
#First turn those NAs into zeros
wtd <- wtd %>%
  mutate(across(c(female, yoy),
                ~ ifelse(is.na(.x) | .x == 0, 0, 1)))

#Next, for event ID lets identify if a young was present or if a female was present
wtd<-wtd %>%
  group_by(Event.ID) %>%
  mutate(
    # Whether any yoy = 1 within the event
    yoy_present = as.integer(any(yoy == 1, na.rm = TRUE)),
    
    # Whether ONLY females were present (female = 1 somewhere, but no yoy = 1 anywhere)
    female_only = as.integer(any(female == 1, na.rm = TRUE) & !any(yoy == 1, na.rm = TRUE))
  ) %>%
  ungroup()

#Finally, lets reduce to one row per event, and filter out events with zeros for both categories (these would represent males and stuff)
wtd<-wtd%>%
  group_by(Event.ID)%>%
  filter(row_number()==1)%>%
  filter(!(yoy_present == 0 & female_only == 0))%>%
  ungroup()%>%
  dplyr::select(-c(female, yoy))


```


Data exploration of independant events
```{r}
#How many ind detections
nrow(wtd)
#How many for each group?
wtd%>%
  #We can sum since they are all 1's
  summarise(n_adult = sum(female_only),
            n_young = sum(yoy_present))

#How many sites?
n_distinct(wtd$site)
#Site occupancy?
n_distinct(wtd$site)/n_distinct(dets$site)

#What do detections look like throughout the year?
#Ok so based on this, a good way to look at young deer would be from June to September each year!
wtd%>%
  mutate(month = month(datetime))%>%
  group_by(month)%>%
  summarise(n_female = sum(female_only),
            n_young = sum(yoy_present))%>%
  pivot_longer(names_to = "demographic_group",
               cols = c("n_female", "n_young"))%>%
  ggplot(aes(x = as.factor(month), y = value, fill = demographic_group))+
  geom_bar(stat = "identity", position = position_dodge())+
  ylab("Independent Detections")+
  ggtitle("Detections of deer ~ month of year")






```

Lets continue to work on our response variable. A couple things need doing. I am going to stretch this into a weekly proportional binomial for each site. Monthly might be too coarse since only using four months. I need to 1) do this properly such that we only count a week as operational if it ran for all 7 days, and then 2) find a way to fill in true zeros
```{r}

#Lets first work on operability of cameras. We will first use the main detection frame as this has all the days of operation for all sites.
#Filter down to june, july, august, and september of each year
df<-dets%>%
  mutate(month = month(datetime),
         date = date(datetime))%>%
  filter(month %in% 6:9)

#Keep only unique dates for each array/site
df<-df%>%
  distinct(array, site, date)

#Identify the year and week for each site, and only keep sites that had full weeks
weekly_operation<-df%>%
  mutate(year = year(date),
         week = week(date))%>%
  #Have to group by year as well, since some sites ran across several years
  group_by(array, year, site, week)%>%
  summarise(n_days = n_distinct(date))%>%
  #Here we remove sites that didn't have full weeks
  filter(n_days == 7)

#How many sites is this? Nice nice nice
n_distinct(weekly_operation$site)
#How many weeks per site? 
weekly_operation%>%
  group_by(year, site)%>%
  summarise(n_weeks = n_distinct(week))%>%
  summarise(mean_weeks = mean(n_weeks),
            sd_weeks = sd(n_weeks),
            min_weeks = min(n_weeks),
            max_weeks = max(n_weeks))

####################################################################################
#Now that we have a list of cameras with operation we can use, I need to identify how many detentions of each group, for each site/year/week combination

#Lets create a temporary frame to work in so we don't have to re-run the independent detection loop again if needed
wtd_dets<-wtd

#First, filter our dataframe to the same months
wtd_dets<-wtd_dets%>%
  mutate(month = month(datetime))%>%
  filter(month %in% 6:9)

#Add the year and week columns we will need
wtd_dets<-wtd_dets%>%
  mutate(year = year(datetime),
         week = week(datetime))

#Ok now I think I can do this. Since each event with a yoy or lone female is indicated by just a 1 (presence), I should be able to sum occurrences for each year/site/week combination right?
#REMEMBER - this will not account for double absences! Sites with zero young and zero adult will not be captured here. Address this next
wtd_dets<-wtd_dets%>%
  group_by(array, year, site, week)%>%
  summarise(sum_yoy = sum(yoy_present),
            sum_adult = sum(female_only))%>%
  #And then take those sums, and turn them back into presence/absence again
  #Remember just want to know for each week if a group was present!
  mutate(across(c(sum_yoy, sum_adult),
                ~ ifelse(is.na(.x) | .x == 0, 0, 1)))%>%
  #rename for clarity
  rename(yoy_present = sum_yoy,
         adult_present = sum_adult)


#And now, we can merge this with the weekly operation file.
wtd_dets<-left_join(weekly_operation, wtd_dets, by = c("array", "year", "site", "week"))

#And now, lets fill in the zeros for the weeks where neither a young or adult was detected
wtd_dets<-wtd_dets%>%
  mutate(yoy_present = ifelse(is.na(yoy_present), 0, yoy_present),
         adult_present = ifelse(is.na(adult_present), 0, adult_present))

#Math check, how many weeks?
wtd_dets%>%
  group_by(year, site)%>%
  summarise(n_weeks = n_distinct(week))%>%
  ungroup()%>%
  summarise(total_weeks = sum(n_weeks))

#And last step, for each array, site, and year --> how many weeks were sampled, and how many of those weeks had a detection of a young or an adult. Again, since young and adult are coded as present or absent, we can sum them up
wtd_dets<-wtd_dets%>%
  group_by(array, year, site)%>%
  summarise(n_weeks = n_distinct(week),   #how many weeks for each site
            n_young = sum(yoy_present),   #how many of those weeks had a young
            n_adult = sum(adult_present)) #how many of those weeks had an adult



#Can we visualize this nicely? Yes, but need to manipulate the frame a little bit for eas
wtd_dets%>%
  pivot_longer(cols = c(n_young, n_adult),
               values_to = "n_present")%>%
ggplot(aes(x = array, y = n_present / n_weeks, color = name))+
  geom_jitter()+
  ylab("Proportion Summer Weeks Present")+
  xlab("Array")+
  scale_color_discrete(
    name = "Age Class",
    labels = c("n_young" = "Fawns", "n_adult" = "Adults"))+
  theme_classic()

```

##############################################################################

Alright lets now work on our covariate file. I am going to keep things very simple for the time being. Just looking at a few 
```{r}
#I am going to reduce this to just a few key players
n_distinct(covs$site)


covs<-covs%>%
  filter(buff_dist == 1000)%>%
  rename(broadleaf = lc_class220)%>%
  relocate(array, site, broadleaf)%>%
  #remove the habitat classifications
  select(-c(lc_class110, lc_class120, lc_class20, lc_class210, lc_class230, lc_class32, lc_class33, lc_class34, lc_class50))

#Create a sum across columns 4 to 50
covs<-covs%>%
  mutate(disturbance_index = rowSums(across(5:50)))%>%
  select(array, site, disturbance_index, broadleaf)


hist(covs$disturbance_index)


#Lets link that into the detection file
wtd_dets<-merge(wtd_dets, covs, by = c("site", "array"))

temp<-wtd_dets
wtd_dets<-temp


```

Weather data. Ok there are a few different things that need doing in this block. Going to keep it all simple and just work with snow for now. But I need to build:
1) RAW: For each site - the preceeding years mean snow data
2) AVERAGE TEMPORAL: for each site, the average mean snow within an array in the preceeding year
3) AVERAGE SPATIAL: for each site, the average mean snow for that site across the last 10 years
4) ANOMALY: for each site the difference between the current data - temporal and - spatial.
```{r}
#So first things first, lets keep life simple and just look at snow
weather_10<-weather_10%>%
  dplyr::select(c(site, latitude, longitude, year, day, snow = snow_water_equiv_kgM2))

#I would like a couple graphs here showing the time series for each variable
#First lets get the dates read in nicely
weather_10 <- weather_10 %>%
  mutate(date = ymd(paste0(year, "-01-01")) + days(day - 1))

weather_summary<-weather_10%>%
  group_by(date)%>%
  summarise(mean_snow = mean(snow),
            sd_snow = sd(snow))

#Snow Plot
ggplot(weather_summary, aes(x = date)) +
  geom_ribbon(
    aes(ymin = mean_snow - sd_snow,
        ymax = mean_snow + sd_snow),
    fill = "grey", alpha = 0.6
  ) +
  geom_line(aes(y = mean_snow), color = "black", size = 1) +
  ylab("Mean Snow Water Equivalent (kg)") +
  xlab("Date") +
  ggtitle("Daily Snow")+
  theme_classic()

######################################################################################
#Ok, lets get these measures into the dataframe. Now, we want to think about lag times and whatnot. So lets extract 

#First lets define the winter months (using guidance from Dickie et al which reports snow periods as Sep 1 to March 31) (Potentially rethink this since we use September data for fawns)
winter_months<-c(9, 10, 11, 12, 1, 2, 3) 


#Ok lets pick an arbitrary winter period. Lets say october to february?
#First just keep data in those months
weather_10<-weather_10%>%
  filter(month(date) %in% winter_months)
  

#Ok ok. Now remember, the weather dataframe has data for all sites for all years. But we don't need that 
weather_summary <- weather_10 %>%
  # Extract month for filtering
  mutate(month = month(date),
         # Define "winter_year" = the year that the winter *starts*
         winter_year = if_else(month >= 9, year(date), year(date) - 1)) %>%
  
  # Group by site and winter season
  group_by(site, winter_year) %>%
  
  # Summarize mean and sd for each variable
  summarise(
    mean_snow  = mean(snow, na.rm = TRUE),
    sd_snow    = sd(snow, na.rm = TRUE),
    n_days    = n(),  # optional — good to check how many days were available
    .groups = "drop"
  )%>%
  mutate(mean_snow = scale(mean_snow))

#Ok so at this point I have the winter_year data for each site. Now I need to manipulate and decompose that data into what I need.

#First, lets get the raw data into the wtd_dets dataframe
#Identify the winter year to use
wtd_dets<-wtd_dets%>%
  mutate(winter_year = year - 1)

wtd_dets<-wtd_dets%>%
  left_join(weather_summary%>%
  dplyr::select(site, winter_year, mean_snow), 
  by = c("site", "winter_year"))

#make sure that worked- no NAs!
summary(wtd_dets$mean_snow)
hist(wtd_dets$mean_snow)

###########################################
#Ok now the temporal component. For each site/year what was the average winter snow across its respective array.
#To do this I will need to add array back into the weather data. My bad. I can get that from camop

weather_summary<-weather_summary%>%
  left_join(camop%>%
  dplyr::select(site, array),
  by = "site")%>%
  relocate(array, site)

#So this calculation is slightly different! Grouping by year to get an overall average!
weather_summary%>%
  group_by(winter_year)%>%
  summarise(mean_snow_temporal = mean(mean_snow))


#join it all up
wtd_dets<-wtd_dets%>%
  left_join(
weather_summary%>%
  group_by(winter_year)%>%
  summarise(mean_snow_temporal = mean(mean_snow)),
by = c("winter_year"))

###################################################################
#Average spatial - so now we are going to take a single site, and find the average over the last 10 years
head(weather_summary)

#This is slightly tricky though, because I need to find a date range for each site. I've got a rather clever idea. Lets take the winter year, and then calculate a min year 10 years before that
site_years<-wtd_dets%>%
  mutate(min_year = winter_year - 10)%>%
  dplyr::select(site, winter_year, min_year)

site_10_year_snow<-site_years %>%
  left_join(weather_summary, by = "site") %>%
  #determining which years to keep based on the winter years in both frames
   #keep weather summary winter year if its greater than the minimum for its site
   filter(winter_year.y >= min_year, 
         #keep weather summary winter year if its less than the max for its site  
         winter_year.y <= winter_year.x) %>%
  group_by(site, winter_year = winter_year.x) %>%
  summarise(mean_snow_spatial = mean(mean_snow, na.rm = TRUE), .groups = "drop")

#And now we can join that in e z p z
wtd_dets<-wtd_dets%>%
  left_join(site_10_year_snow, by = c("site", "winter_year"))

############################################################################
#Winter anomaly!

#This one should be easier I think to calculate as I am pretty sure it is just the difference between the raw and the time minus space
wtd_dets<-wtd_dets%>%
  mutate(snow_anomaly = mean_snow - mean_snow_temporal - mean_snow_spatial)

hist(wtd_dets$snow_anomaly)
hist(wtd_dets$mean_snow)
hist(wtd_dets$mean_snow_temporal)
hist(wtd_dets$mean_snow_spatial)




```


Lets do some quick model prep and modeling before lunch break
```{r}
#Lets scale the remaining predictors
wtd_dets<-wtd_dets%>%
  mutate(disturbance = scale(disturbance_index))

#How do I keep losing lat/long??
wtd_dets<-wtd_dets%>%
  left_join(camop%>%
              dplyr::select(site, lat, long),
            by = "site")

covs

#I am sure things are correlated af but I dont care right now
# Select your variables
vars <- wtd_dets %>%
  dplyr::select(mean_snow, mean_snow_temporal, mean_snow_spatial, snow_anomaly, disturbance, broadleaf)

# Compute correlation matrix (use complete.obs to ignore NAs)
cor_matrix <- cor(vars, use = "complete.obs")

# Print the correlation matrix
print(round(cor_matrix, 2))

# Visualize it
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, addCoef.col = "black")
```

Adult models
```{r}
#Who cares lets build
disturbance_a<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   disturbance +scale(lat) + broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_a)
snow_a<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   mean_snow +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(snow_a)

temporal_a<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   mean_snow_temporal + scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(temporal_a)

spatial_a<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   mean_snow_spatial + scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(spatial_a)


anomaly_a<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   snow_anomaly +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(anomaly_a)

snow_disturbance_a<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   mean_snow*disturbance +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(snow_disturbance_a)

temporal_disturbance_a<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   mean_snow_temporal*disturbance +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(temporal_disturbance_a)

spatial_disturbance_a<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   mean_snow_spatial*disturbance +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(spatial_disturbance_a)


anomaly_disturbance_a<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   snow_anomaly*disturbance +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(anomaly_disturbance_a)

library(MuMIn)
ms_a<-model.sel(disturbance_a, snow_a, temporal_a, spatial_a, anomaly_a,
          snow_disturbance_a, temporal_disturbance_a, spatial_disturbance_a, anomaly_disturbance_a,
          rank = "AICc")

ms_a<- as.data.frame(ms_a) %>%
  # Add model names as a column
  mutate(model = rownames(ms_a)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

View(ms_a)

summary(spatial_a)


```


Young
```{r}
#Who cares lets build
disturbance_y<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   disturbance +scale(lat) + broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_y)
snow_y<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   mean_snow +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(snow_y)

temporal_y<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   mean_snow_temporal + scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(temporal_y)

spatial_y<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   mean_snow_spatial + scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(spatial_y)


anomaly_y<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   snow_anomaly +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(anomaly_y)

snow_disturbance_y<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   mean_snow*disturbance +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(snow_disturbance_y)

temporal_disturbance_y<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   mean_snow_temporal*disturbance +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(temporal_disturbance_y)

spatial_disturbance_y<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   mean_snow_spatial*disturbance +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(spatial_disturbance_y)


anomaly_disturbance_y<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   snow_anomaly*disturbance +scale(lat) +broadleaf +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(anomaly_disturbance_y)

library(MuMIn)
ms_y<-model.sel(disturbance_y, snow_y, temporal_y, spatial_y, anomaly_y,
          snow_disturbance_y, temporal_disturbance_y, spatial_disturbance_y, anomaly_disturbance_y,
          rank = "AICc")

ms_y<- as.data.frame(ms_y) %>%
  # Add model names as a column
  mutate(model = rownames(ms_y)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

View(ms_y)


```


```




Preparing the data for preliminary models
```{r}
#Lets scale the variables and then explore correlations
temp<-temp%>%
  mutate(across(
    .cols = c(broadleaf, roads, wells, conventional_seismic, pipeline, mean_min_temp, mean_swe, mean_precip, snow_dif, temp_dif, precip_dif, snow_10_mean),
    .fns = ~ as.numeric(scale(.x)),
    .names = "scale_{.col}"
  ))

#Check for correlations
# Select your variables
vars <- temp %>%
  dplyr::select(scale_broadleaf, scale_roads, scale_wells, scale_conventional_seismic, scale_pipeline, scale_mean_min_temp, scale_mean_swe, scale_mean_precip, snow_dif, temp_dif, precip_dif, snow_10_mean)

# Compute correlation matrix (use complete.obs to ignore NAs)
cor_matrix <- cor(vars, use = "complete.obs")

# Print the correlation matrix
print(round(cor_matrix, 2))

# Visualize it
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, addCoef.col = "black")

#Alright precipitation is a bad guy. lets remove it
temp<-temp%>%
  dplyr::select(-c(scale_mean_precip, precip_df)


```


Its all come to down to this. Lets run some models baby
```{r}

#Variance estimates for year were super low, so not including those as a random effect

#global model for adults and young to get started!
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_mean_min_temp + scale_mean_swe +
                   (1|array) + (1|year),
                 family = "binomial",
                 data = temp)
summary(m1)


m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_mean_min_temp + scale_mean_swe +
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)

####################
#interaction terms?
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_mean_min_temp + scale_mean_swe + scale_roads*scale_mean_min_temp +
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)


m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_mean_min_temp + scale_mean_swe + scale_pipeline*scale_mean_min_temp+
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)

###################################
#What about a nice straightforward interaction?
m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_wells*scale_mean_swe+
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)


#What if we account for that north south gradient?
#Adding the coordinates back in
temp<-temp%>%
  left_join(camop%>%
  dplyr::select(site, lat, long),
  by = c("site"))

#Ok yea latitude is important for suresies
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf + lat +
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf + lat + scale_mean_swe +
                       scale_pipeline+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


#Models with the difference stuff
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf + scale_snow_10_mean +
                       scale_pipeline+ scale_snow_dif+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


#Models with the difference stuff
m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf + scale_snow_10_mean + scale_snow_dif +
                       scale_pipeline+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf + scale_snow_10_mean +
                       scale_roads+ scale_snow_dif+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


#Models with the difference stuff
m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf + scale_snow_10_mean + scale_snow_dif +
                       scale_pipeline+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


#Global
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_snow_10_mean + scale_snow_dif +
                   (1|array) + (1|year),
                 family = "binomial",
                 data = temp)
summary(m1)

#Global
m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_snow_10_mean + scale_snow_dif +
                   (1|array) + (1|year),
                 family = "binomial",
                 data = temp)
summary(m1)



#Global with interaction
m1_adult<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_snow_10_mean + scale_snow_dif + scale_snow_10_mean*scale_roads + scale_snow_dif*scale_roads+
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)

#Global with interaction
m1_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_snow_10_mean + scale_snow_dif +scale_snow_10_mean*scale_roads + scale_snow_dif*scale_roads+
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)

# -----------------------------
# 1. Predicted probabilities for adult-only model
# -----------------------------
roads_q <- quantile(temp$scale_roads, probs = c(0.25, 0.75), na.rm = TRUE)
low_roads <- roads_q[1]
high_roads <- roads_q[2]


adult_pred_low <- ggpredict(m1_adult, terms = c("scale_snow_10_mean [all]", paste0("scale_roads [", round(low_roads, 2), "]")))
adult_pred_low$road_level <- "Low Roads"

adult_pred_high <- ggpredict(m1_adult, terms = c("scale_snow_10_mean [all]", paste0("scale_roads [", round(high_roads, 2), "]")))
adult_pred_high$road_level <- "High Roads"

adult_pred <- bind_rows(adult_pred_low, adult_pred_high)
adult_pred$group <- "Adult Female"

# -----------------------------
# 3. Predicted probabilities for adult+young model
# -----------------------------
young_pred_low <- ggpredict(m1_young, terms = c("scale_snow_10_mean [all]", paste0("scale_roads [", round(low_roads, 2), "]")))
young_pred_low$road_level <- "Low Roads"

young_pred_high <- ggpredict(m1_young, terms = c("scale_snow_10_mean [all]", paste0("scale_roads [", round(high_roads, 2), "]")))
young_pred_high$road_level <- "High Roads"

young_pred <- bind_rows(young_pred_low, young_pred_high)
young_pred$group <- "Adult + Young"

pred_df <- bind_rows(adult_pred, young_pred)

ggplot(pred_df, aes(x = x, y = predicted, color = road_level)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = road_level), alpha = 0.2, color = NA) +
  facet_wrap(~group) +
  labs(
    x = "Long-term Snow Load (10-year mean, scaled)",
    y = "Predicted Probability of Detection",
    color = "Road Density",
    fill = "Road Density"
  ) +
  scale_color_manual(values = c("blue", "red")) +
  scale_fill_manual(values = c("blue", "red")) +
  theme_minimal(base_size = 14)
```
