This to-do in this section
1. Code models for each demographic group - polygonal and linear
2. Model predictions for each group - 3 main figures

```{r}
#Clear everything out and start fresh
rm(list=ls())
library(reproducible) #setting up directories for data management
library(readr)        #Needed to read and merge all the csv's in the camera folder
library(dplyr)        #data wranglin'
library(lubridate)    #datetime manipulation
library(tidyr)        #more data manipulation
library(stringr)      #detecting character strings
library(ggplot2)      #data visualization
library(purrr)        #feline-friendly loops 
library(corrplot)     #correlation tests
library(MuMIn)
library(ggeffects)


#This just sets up files for inputs (already created, must contain your files), and outputs
input_directory<-reproducible::checkPath(file.path(getwd(), "3. Model Selection/outputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "4. Models/outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "4. Models/figures"), create = TRUE)


#read in the files we need
#This is from the species response variable file
wtd_dets<-read.csv(file.path(input_directory, "final_df.csv"))



```

Ok everything has been prepared for modeling, so I guess I can just get to it. So a few things to do for each demographic group. Fit the linear and the polygonal model, report the coefficients, get the model predictions, run the diagnostics. Wondering if the diagnostics should maybe be in another script but we will see 

This is a silly little chunk, but I need to define the linear and polygonal lower and upper quantiles that we will make model predictions for. Rather than randomly do this in one of the chunks, I am just doing it here (only need to do it once since the values to predict for are all the same!)
```{r}
#Lower and upper quantiles for linear
linear_QL<-quantile(wtd_dets$linear_index, 0.10, na.rm = TRUE)
linear_QH<-quantile(wtd_dets$linear_index, 0.90, na.rm = TRUE)

#Lower and upper quantiles for polygonal
polygonal_QL<-quantile(wtd_dets$polygonal_index, 0.10, na.rm = TRUE)
polygonal_QH<-quantile(wtd_dets$polygonal_index, 0.90, na.rm = TRUE)


```



Males
```{r}
#1. Fit the models
#2.  Model Predictions


#Linear model
linear_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_males)

#


#Polygonal model
polygonal_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_males)




##############################################################################
#First, LRT tests for each model


```


Now for males
```{r}

#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_males)

#What about a "null" model that doesn't include disturbance?
null_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_males)



#All disturbance
# disturbance_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
#                    disturbance_index+
#                    PC1 + PC2 + #Habitat covariates
#                    predator_index + #Predator index
#                    lat+
#                    mean_snow_spatial*disturbance_index+
#                    snow_anomaly*disturbance_index+
#                    (1|array),
#                  family = "binomial",
#                  data = wtd_dets)
# summary(disturbance_males)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_males)


#Linear
linear_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_males)



#Model selection and cleanup into a nice dataframe!
model_selection_male<- as.data.frame(model.sel(global_males, null_males, linear_males, polygonal_males,
              rank = "AICc"))

model_selection_male<-model_selection_male%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_male)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_male)<-NULL




```


Females
```{r}
#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_females)

#What about a "null" model that doesn't include disturbance?
null_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_females)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_females)


#Linear
linear_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_females)



#Model selection and cleanup into a nice dataframe!
model_selection_females<- as.data.frame(model.sel(global_females, null_females, linear_females, polygonal_females,
              rank = "AICc"))

model_selection_females<-model_selection_females%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_females)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_females)<-NULL




```


Young Deer?
```{r}
#All disturbance
#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_young)

#What about a "null" model that doesn't include disturbance?
null_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_young)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_young)


#Linear
linear_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_young)



#Model selection and cleanup into a nice dataframe!
model_selection_young<- as.data.frame(model.sel(global_young, null_young, linear_young, polygonal_young,
              rank = "AICc"))

model_selection_young<-model_selection_young%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_young)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_young)<-NULL



```


Ok, in the following here is what I want to do. I want to write a loop that performs an out of sample test of our model selection, to identify where the "best" top model is identified
```{r}

#These are the response variables to iterate over
response_variables<-data.frame(response = c("cbind(n_male, (n_weeks - n_male)) ~",
                      "cbind(n_female, (n_weeks - n_female)) ~",
                      "cbind(n_young, (n_weeks - n_young)) ~"),
                      name = c("males", "females", "young"))


#Create a blank list of model selection tables
model_selection_tables<-list()

#How many simulations to do?
n_sims<-100



for(j in 1:nrow(response_variables)){
    
  #Store the current response variable we are working on 
  response <- as.character(response_variables[j, 1])
  #And the name, we will need that for later
  name<- as.character(response_variables[j, 2])
  
  print(paste("Working on models for", name, "at", Sys.time()))
  
  
  #Pre-emptive, but its more efficient to define the model formulas here within this loop, so its only done once per group rather than everytime within the next loop! #Efficiency 101 baby
  formula_global<- as.formula(paste(response,
                      "polygonal_index+linear_index+PC1 + PC2 +predator_index +lat+mean_snow_spatial+snow_anomaly+
                           + (1|array)", sep = ""))
  
  formula_null<- as.formula(paste(response,
                      "PC1 + PC2 +predator_index +lat+mean_snow_spatial+snow_anomaly+
                           + (1|array)", sep = "")) 
  formula_linear<- as.formula(paste(response,
                      "linear_index+
                      PC1 + PC2 +predator_index +lat+
                      mean_snow_spatial*linear_index+
                      snow_anomaly*linear_index+
                           + (1|array)", sep = ""))
  formula_polygonal<- as.formula(paste(response,
                      "polygonal_index+
                      PC1 + PC2 +predator_index +lat+
                      mean_snow_spatial*polygonal_index+
                      snow_anomaly*polygonal_index+
                           + (1|array)", sep = ""))
  
  
  #Now that we have identified our response variable, we need to many
  #many times sample the data, fit each model, run model selection,
  #and identify the top model. 
  for(i in 1:n_sims){
    
    #Sample the dataframe - THIS IS WHERE WE DEFINE THE SAMPLING AMOUNT
    sample<-wtd_dets%>%
      slice_sample(prop = 0.90)
    
    
    #Fit each model!
    fit_global<-glmmTMB::glmmTMB(
                formula = formula_global,
                 family = "binomial",
                 data = sample)
    
    fit_null<-glmmTMB::glmmTMB(
                formula = formula_null,
                 family = "binomial",
                 data = sample)
    
    fit_linear<-glmmTMB::glmmTMB(
                formula = formula_linear,
                 family = "binomial",
                 data = sample)
    
    fit_polygonal<-glmmTMB::glmmTMB(
                formula = formula_polygonal,
                 family = "binomial",
                 data = sample)
    
    #Quick model selection! and cleanup
   model_sel<-as.data.frame(MuMIn::model.sel(fit_global, fit_null, fit_linear, fit_polygonal,
              rank = "AICc"))
   
   model_sel<-model_sel%>%
    # Add model names as a column
  mutate(model = rownames(model_sel)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))
   
   #Get rid of the rownames, they are annoying
   rownames(model_sel)<-NULL
   
   #Add the name, and the simulation number to this dataframe
   model_sel<-model_sel%>%
     mutate(group = name,
            simulation = i)%>%
     relocate(group, simulation)
   
   # Initialize the nested list if this is the first simulation for this group
if (! name %in% names(model_selection_tables)) {
  model_selection_tables[[name]] <- list()
}

# Store the model table for this simulation
model_selection_tables[[name]][[i]] <- model_sel
  }
  
  #Nothing to see here!
  
}

#Ok now lets take those results and strip them out into a single dataframe
all_model_results <- purrr::map_df(
  model_selection_tables,
  ~ purrr::map_df(.x, ~ .x)
)

#And finally, for each group, how often was the top model top?
all_models <- c("fit_linear", "fit_polygonal", "fit_global", "fit_null")


summary_table <-
  all_model_results %>%
  # Only the top model per simulation
  filter(delta == 0) %>%
  group_by(group, model) %>%
  summarise(n_obs = n(), .groups = "drop") %>%
  
  # Expand so every group has all 4 models
  tidyr::complete(
    group,
    model = all_models,
    fill = list(n_obs = 0)
  ) %>%
  
  # Add total sims and proportion
  mutate(
    total = n_sims,
    prop_top = n_obs / total
  )

#And a plot!
# Set factor order
summary_table <- summary_table %>%
  mutate(
    group = factor(group, levels = c("males", "females", "young")),
    model = factor(model,
                   levels = c("fit_linear", "fit_polygonal", "fit_global", "fit_null"),
                   labels = c("Linear", "Polygonal", "Global", "Null"))
  )

ggplot(summary_table, aes(x = model, y = prop_top, fill = model)) +
  geom_col(width = 0.7, alpha = 0.9) +
  facet_wrap(~ group) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     limits = c(0, 1)) +
  labs(
    title = "Proportion of Simulations in Which Each Model Was Top",
    x = "Model",
    y = "Proportion Top Model"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )

```
