This to-do in this section
1. Code models for each demographic group - polygonal and linear
2. Model predictions for each group - 3 main figures

```{r}
#Clear everything out and start fresh
rm(list=ls())
library(reproducible) #setting up directories for data management
library(readr)        #Needed to read and merge all the csv's in the camera folder
library(dplyr)        #data wranglin'
library(lubridate)    #datetime manipulation
library(tidyr)        #more data manipulation
library(stringr)      #detecting character strings
library(ggplot2)      #data visualization
library(purrr)        #feline-friendly loops 
library(corrplot)     #correlation tests

library(performance)  #Model Validation
library(ggeffects)    #Model predictions
library(broom.mixed)  #Extracting model coefficients
library(flextable)    #Formatting and writing to worddoc
library(officer)      #Formatting and writing to wordoc


#This just sets up files for inputs (already created, must contain your files), and outputs
input_directory<-reproducible::checkPath(file.path(getwd(), "3. Model Selection/outputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "4. Models/outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "4. Models/figures"), create = TRUE)


#read in the files we need
#This is from the species response variable file
wtd_dets<-read.csv(file.path(input_directory, "final_df.csv"))



```


Fit the models and extract coefficient estimates
```{r}
linear_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_males)

polygonal_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_males)

linear_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_females)

polygonal_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_females)

linear_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_young)

polygonal_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_young)

#Lets save those six models to the output directory, we will need them for the next steps of simulation
saveRDS(linear_males, file = file.path(output_directory, "linear_males.rds"))
saveRDS(linear_females, file = file.path(output_directory, "linear_females.rds"))
saveRDS(linear_young, file = file.path(output_directory, "linear_young.rds"))
saveRDS(polygonal_males, file = file.path(output_directory, "polygonal_males.rds"))
saveRDS(polygonal_females, file = file.path(output_directory, "polygonal_females.rds"))
saveRDS(polygonal_young, file = file.path(output_directory, "polygonal_young.rds"))


#Alright now that we have our six models, can we extract model coefficients in an efficient way?
models<-list(
  linear_males      = linear_males,
  polygonal_males   = polygonal_males,
  linear_females    = linear_females,
  polygonal_females = polygonal_females,
  linear_young      = linear_young,
  polygonal_young   = polygonal_young
)

#Whip up a funciton that extracts and reshapes
extract_coefficients<-function(model){
  
  broom.mixed::tidy(model)
  
}


broom.mixed::tidy(linear_females)

df<-purrr::imap_dfr(models, ~{
  extract_coefficients(.x)%>%
    mutate(model_name = .y)
  
})

#Lets clean this up a little bit 
df<-df%>%
  dplyr::select(term, estimate, std.error, p.value, model_name)%>%
  relocate(model_name)

#Recode those couple terms
df <- df %>%
  mutate(term = dplyr::case_when(
    term == "(Intercept)"      ~ "Intercept",
    term == "sd__(Intercept)"  ~ "Array ID",
    TRUE ~ term
  ))

#Quick funciton to format P values
format_p <- function(p) {
  case_when(
    is.na(p) ~ NA_character_,
    p < 0.001 ~ "P < 0.001",
    p < 0.01  ~ "P < 0.01",
    p < 0.05  ~ "P < 0.05",
    TRUE ~ paste0("P = ", round(p, 3))
  )
}


df <- df %>%
  mutate(
    estimate_round = round(estimate, 2),
    se_round       = round(std.error, 2),
    p_formatted    = format_p(p.value),
    estimate = paste0(
      estimate_round, " \u00B1 ", se_round, ", ", p_formatted
    )
  ) %>%
  select(model_name, term, estimate)


df_wide <- df %>%
  tidyr::pivot_wider(
    names_from = model_name,
    values_from = estimate
  )

#And now, re-order values and arrange!
df_wide$term<-factor(df_wide$term, 
                     levels = c("Intercept", "PC1", "PC2", "predator_index", "lat", "mean_snow_spatial", "snow_anomaly", "linear_index","linear_index:mean_snow_spatial", "linear_index:snow_anomaly", "polygonal_index", "polygonal_index:mean_snow_spatial", "polygonal_index:snow_anomaly", "Array ID"))

df_wide<-df_wide%>%
  arrange(term)

#And then finally, format and shape that into a word document table
# Convert to a nice Word table
ft <- flextable::flextable(df_wide)

# Optimize the widths and heights of columns. Nice to have!
ft <- flextable::autofit(ft)

#And then turn that into an actual word document
doc <- officer::read_docx()
doc <- flextable::body_add_flextable(doc, ft)

#Now actually write the file
#NOTE - this table is quite big so you'll have to modify it into landscape style manually
#I tried to do it in the code, but got lost
print(doc, target = file.path(output_directory, "model_coefficients.docx"))

```

Model diagnostics and whatnot go here
```{r}
#Things to check in here
#Variance Inflation Factors
#Spatial Autocorrelation  Morans I
#Random effect LRT
#

#Variance Inflation Factors for all models
#We can use the list of models defined above for this
models

#Create a function to extract the VIFs
extract_vif <- function(model) {
  vif_obj <- performance::check_collinearity(model)
  as.data.frame(vif_obj)[, c("Term", "VIF")]
}

#And then use that function to create a dataframe of VIFs for each model object
vif_df_long <- map_df(
  names(models),
  ~ extract_vif(models[[.x]]) %>%
    mutate(model = .x),
  .id = NULL
)

#Slightly re-arrange that dataframe
vif_df_long<-vif_df_long%>%
  mutate(VIF = round(VIF, digits = 2))%>%
  pivot_wider(names_from = model, values_from = VIF)

#Export as a csv file
write.csv(vif_df_long, file.path(output_directory, "variance_inflation_factors.csv"),
          row.names = FALSE)

# Convert to word doc table for export
ft <- flextable::flextable(vif_df_long)

# Optimize the widths and heights of columns. Nice to have!
ft <- flextable::autofit(ft)

#And then turn that into an actual word document
doc <- officer::read_docx()
doc <- flextable::body_add_flextable(doc, ft)

#Now actually write the file
#NOTE - this table is quite big so you'll have to modify it into landscape style manually
#I tried to do it in the code, but got lost
print(doc, target = file.path(output_directory, "variance_inflation_factors.docx"))

#########################################################################################
#Measures of Model Fit
#Same thing, we can use the list of models, and write a quick function to extract all these values

extract_r2<-function(model){
  model_fit<-performance::r2(model)
  as.data.frame(model_fit)[, c("R2_conditional", "R2_marginal")]
}

 model_r2<-map_df(
  names(models),
  ~ extract_r2(models[[.x]]) %>%
    mutate(model = .x),
  .id = NULL
)%>%
   relocate(model)

#Write that csv
write.csv(model_r2, file.path(output_directory, "model_r2.csv"),
          row.names = FALSE)
  

#Likelihood Ratio Tests for random effects
#This is a bit more indepth because I need to fit the same models above, just without the random effect...
linear_males_reduced<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index,
                 family = "binomial",
                 data = wtd_dets)



polygonal_males_reduced<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index,
                 family = "binomial",
                 data = wtd_dets)


linear_females_reduced<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index,
                   
                 family = "binomial",
                 data = wtd_dets)

polygonal_females_reduced<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index,
                 family = "binomial",
                 data = wtd_dets)


linear_young_reduced<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index,
                 family = "binomial",
                 data = wtd_dets)


polygonal_young_reduced<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index,
                 family = "binomial",
                 data = wtd_dets)

#Ok, create a list of the paired full and reducted models
lrt_models <- list(
  linear_males      = list(full = linear_males,      reduced = linear_males_reduced),
  polygonal_males   = list(full = polygonal_males,   reduced = polygonal_males_reduced),
  linear_females    = list(full = linear_females,    reduced = linear_females_reduced),
  polygonal_females = list(full = polygonal_females, reduced = polygonal_females_reduced),
  linear_young      = list(full = linear_young,      reduced = linear_young_reduced),
  polygonal_young   = list(full = polygonal_young,   reduced = polygonal_young_reduced)
)

#And yet again, make a quick little function
run_lrt <- function(full, reduced) {
  out <- lmtest::lrtest(full, reduced)
  broom::tidy(out)  # makes it easy to combine
}

#Then run that function across the full and reduced models, spitting out a dataframe
#Its going to give errors but thats fine, the LRT dataframe is funky because it misses entries for some rows
lrt_results <- map_df(
  names(lrt_models),
  ~ run_lrt(lrt_models[[.x]]$reduced, lrt_models[[.x]]$full) %>%
      mutate(model_name = .x)
)

#But now we can take that dataframe and extract just what we need for each model (test statistic, df, and P value)
lrt_results<-lrt_results%>%
  dplyr::select(model_name, df , statistic, p.value)%>%
  #Then to be lazy, just keep df = 1 (only need one obs per model name)
  filter(df == 1)

#And write that file
write.csv(lrt_results, file.path(output_directory, "lrt_results.csv"),
          row.names = FALSE)
```

Should we explore spatial autocorrelation?
```{r}




```




Ok everything has been prepared for modeling, so I guess I can just get to it. So a few things to do for each demographic group. Fit the linear and the polygonal model, report the coefficients, get the model predictions, run the diagnostics. Wondering if the diagnostics should maybe be in another script but we will see 

This is a silly little chunk, but I need to define the linear and polygonal lower and upper quantiles that we will make model predictions for. Rather than randomly do this in one of the chunks, I am just doing it here (only need to do it once since the values to predict for are all the same!)
```{r}
#Lower and upper quantiles for linear
linear_QL<-quantile(wtd_dets$linear_index, 0.10, na.rm = TRUE)
linear_QH<-quantile(wtd_dets$linear_index, 0.90, na.rm = TRUE)

#Lower and upper quantiles for polygonal
polygonal_QL<-quantile(wtd_dets$polygonal_index, 0.10, na.rm = TRUE)
polygonal_QH<-quantile(wtd_dets$polygonal_index, 0.90, na.rm = TRUE)

#Define the colour blind paletee 
#Linear plot
okabe_ito_linear <- c(
  "10%" = "#0072B2",   # blue
  "90%" = "#D55E00"    # vermillion
)
okabe_ito_polgyonal <- c(
  "10%" = "#009E73",   # blue
  "90%" = "#E69F00"    # vermillion
)
#And this is fun, so in the model selection script, I scaled all the data for modeling process, but later when I make model predictions I want predictions for the unscaled data. So I am going to temporarily read an old file (before things were scaled), extract the information I need (mean and sd for relevant variables), and then remove that dataframe. Savy?
temp_data<-read.csv(file.path(getwd(), "2. Site Variables/outputs/final_df.csv"))

#long term snow
snow_spatial_mean<-mean(temp_data$mean_snow_spatial)
snow_spatial_sd<-sd(temp_data$mean_snow_spatial)

#Snow anomaly
snow_anomaly_mean<-mean(temp_data$snow_anomaly)
snow_anomaly_sd<-sd(temp_data$snow_anomaly)


#And get rid of that dusty old dataframe!
rm(temp_data)

male_axis_limits<-c(0, 0.40)
female_axis_limits<-c(0, 0.40)
young_axis_limits<-c(0, 0.35)

```



Males
```{r}
#1. Fit the models
#2.  Model Predictions


#Linear model
linear_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_males)

#Make predictions from that model
preds_lms<-data.frame(ggeffects::ggpredict(
             linear_males,
  terms = c("mean_snow_spatial [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         linear_quantile = group))

#Back transform the snow variable
preds_lms<- preds_lms %>%
  mutate( snow_spatial_unscaled = snow_spatial * snow_spatial_sd + snow_spatial_mean)


# Ensure your quantile variable has correct labels
preds_lms$linear_quantile_label <- dplyr::recode(preds_lms$linear_quantile,
                                                "-1.13"  = "10%",
                                                "1.04" = "90%")
#And then the actual plot
linear_male_snow<-ggplot(preds_lms, aes(x = snow_spatial_unscaled,
                     y = predicted,
                     colour = linear_quantile_label)) +
  geom_line(size = 1.2) +
  ylim(male_axis_limits)+
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = linear_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_linear,
                      name = "Linear Index Quantile") +
  scale_fill_manual(values = okabe_ito_linear,
                    name = "Linear Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("10 Year Average Winter\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)

########################################################################################
#And now predictions for the Snow Anomaly
preds_lma<-data.frame(ggeffects::ggpredict(
             linear_males,
  terms = c("snow_anomaly [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         linear_quantile = group))

#Back transform the snow variable
preds_lma<- preds_lma %>%
  mutate( snow_anomaly_unscaled = snow_anomaly * snow_anomaly_sd + snow_anomaly_mean)

hist(preds_lma$snow_anomaly_unscaled)

# Ensure your quantile variable has correct labels
preds_lma$linear_quantile_label <- dplyr::recode(preds_lma$linear_quantile,
                                                "-1.13"  = "10%",
                                                "1.04" = "90%")
#And then the actual plot
linear_male_anomaly<-ggplot(preds_lma, aes(x = snow_anomaly_unscaled,
                     y = predicted,
                     colour = linear_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = linear_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_linear,
                      name = "Linear Index Quantile") +
  scale_fill_manual(values = okabe_ito_linear,
                    name = "Linear Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("Snow Anomaly\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(male_axis_limits)

########################################################################################
########################################################################################
########################################################################################
#Polygonal model and predictions
polygonal_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_males)


#Make predictions from that model
preds_pms<-data.frame(ggeffects::ggpredict(
             polygonal_males,
  terms = c("mean_snow_spatial [all]",
            paste0("polygonal_index [", round(polygonal_QL, 2), ",", round(polygonal_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         polygonal_quantile = group))

#Back transform the snow variable
preds_pms<- preds_pms %>%
  mutate( snow_spatial_unscaled = snow_spatial * snow_spatial_sd + snow_spatial_mean)

hist(preds_pms$snow_spatial_unscaled)

# Ensure your quantile variable has correct labels
preds_pms$polygonal_quantile_label <- dplyr::recode(preds_pms$polygonal_quantile,
                                                "-0.93"  = "10%",
                                                "1.39" = "90%")
#And then the actual plot
polygonal_male_snow<-ggplot(preds_pms, aes(x = snow_spatial_unscaled,
                     y = predicted,
                     colour = polygonal_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = polygonal_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_polgyonal,
                      name = "Polygonal Index Quantile") +
  scale_fill_manual(values = okabe_ito_polgyonal,
                    name = "Polygonal Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("10 Year Average Winter\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(male_axis_limits)

#And now the anomaly plots
#Polygonal model and predictions


#Make predictions from that model
preds_pma<-data.frame(ggeffects::ggpredict(
             polygonal_males,
  terms = c("snow_anomaly [all]",
            paste0("polygonal_index [", round(polygonal_QL, 2), ",", round(polygonal_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         polygonal_quantile = group))

#Back transform the snow variable
preds_pma<- preds_pma %>%
  mutate( snow_anomaly_unscaled = snow_anomaly * snow_anomaly_sd + snow_anomaly_mean)

hist(preds_pma$snow_anomaly_unscaled)

# Ensure your quantile variable has correct labels
preds_pma$polygonal_quantile_label <- dplyr::recode(preds_pma$polygonal_quantile,
                                                "-0.93"  = "10%",
                                                "1.39" = "90%")
#And then the actual plot
polygonal_male_anomaly<-ggplot(preds_pma, aes(x = snow_anomaly_unscaled,
                     y = predicted,
                     colour = polygonal_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = polygonal_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_polgyonal,
                      name = "Polygonal Index Quantile") +
  scale_fill_manual(values = okabe_ito_polgyonal,
                    name = "Polygonal Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("Snow Anomaly\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(male_axis_limits)

```

Females
```{r}


#Linear model
linear_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_females)

#Make predictions from that model
preds_lfs<-data.frame(ggeffects::ggpredict(
             linear_females,
  terms = c("mean_snow_spatial [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         linear_quantile = group))

#Back transform the snow variable
preds_lfs<- preds_lfs %>%
  mutate( snow_spatial_unscaled = snow_spatial * snow_spatial_sd + snow_spatial_mean)


# Ensure your quantile variable has correct labels
preds_lfs$linear_quantile_label <- dplyr::recode(preds_lfs$linear_quantile,
                                                "-1.13"  = "10%",
                                                "1.04" = "90%")
#And then the actual plot
linear_female_snow<-ggplot(preds_lfs, aes(x = snow_spatial_unscaled,
                     y = predicted,
                     colour = linear_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = linear_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_linear,
                      name = "Linear Index Quantile") +
  scale_fill_manual(values = okabe_ito_linear,
                    name = "Linear Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("10 Year Average Winter\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(female_axis_limits)

########################################################################################
#And now predictions for the Snow Anomaly
preds_lfa<-data.frame(ggeffects::ggpredict(
             linear_females,
  terms = c("snow_anomaly [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         linear_quantile = group))

#Back transform the snow variable
preds_lfa<- preds_lfa %>%
  mutate( snow_anomaly_unscaled = snow_anomaly * snow_anomaly_sd + snow_anomaly_mean)

hist(preds_lfa$snow_anomaly_unscaled)

# Ensure your quantile variable has correct labels
preds_lfa$linear_quantile_label <- dplyr::recode(preds_lfa$linear_quantile,
                                                "-1.13"  = "10%",
                                                "1.04" = "90%")
#And then the actual plot
linear_female_anomaly<-ggplot(preds_lfa, aes(x = snow_anomaly_unscaled,
                     y = predicted,
                     colour = linear_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = linear_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_linear,
                      name = "Linear Index Quantile") +
  scale_fill_manual(values = okabe_ito_linear,
                    name = "Linear Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("Snow Anomaly\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(female_axis_limits)

########################################################################################
########################################################################################
########################################################################################
#Polygonal model and predictions
polygonal_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_females)


#Make predictions from that model
preds_pfs<-data.frame(ggeffects::ggpredict(
             polygonal_females,
  terms = c("mean_snow_spatial [all]",
            paste0("polygonal_index [", round(polygonal_QL, 2), ",", round(polygonal_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         polygonal_quantile = group))

#Back transform the snow variable
preds_pfs<- preds_pfs %>%
  mutate( snow_spatial_unscaled = snow_spatial * snow_spatial_sd + snow_spatial_mean)

hist(preds_pfs$snow_spatial_unscaled)

# Ensure your quantile variable has correct labels
preds_pfs$polygonal_quantile_label <- dplyr::recode(preds_pfs$polygonal_quantile,
                                                "-0.93"  = "10%",
                                                "1.39" = "90%")
#And then the actual plot
polygonal_female_snow<-ggplot(preds_pfs, aes(x = snow_spatial_unscaled,
                     y = predicted,
                     colour = polygonal_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = polygonal_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_polgyonal,
                      name = "Polygonal Index Quantile") +
  scale_fill_manual(values = okabe_ito_polgyonal,
                    name = "Polygonal Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("10 Year Average Winter\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(female_axis_limits)

#And now the anomaly plots
#Polygonal model and predictions


#Make predictions from that model
preds_pfa<-data.frame(ggeffects::ggpredict(
             polygonal_females,
  terms = c("snow_anomaly [all]",
            paste0("polygonal_index [", round(polygonal_QL, 2), ",", round(polygonal_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         polygonal_quantile = group))

#Back transform the snow variable
preds_pfa<- preds_pfa %>%
  mutate( snow_anomaly_unscaled = snow_anomaly * snow_anomaly_sd + snow_anomaly_mean)

hist(preds_pfa$snow_anomaly_unscaled)

# Ensure your quantile variable has correct labels
preds_pfa$polygonal_quantile_label <- dplyr::recode(preds_pfa$polygonal_quantile,
                                                "-0.93"  = "10%",
                                                "1.39" = "90%")
#And then the actual plot
polygonal_female_anomaly<-ggplot(preds_pfa, aes(x = snow_anomaly_unscaled,
                     y = predicted,
                     colour = polygonal_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = polygonal_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_polgyonal,
                      name = "Polygonal Index Quantile") +
  scale_fill_manual(values = okabe_ito_polgyonal,
                    name = "Polygonal Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("Snow Anomaly\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(female_axis_limits)

```


Young
```{r}
#1. Fit the models
#2.  Model Predictions


#Linear model
linear_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_young)

#Make predictions from that model
preds_lys<-data.frame(ggeffects::ggpredict(
             linear_young,
  terms = c("mean_snow_spatial [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         linear_quantile = group))

#Back transform the snow variable
preds_lys<- preds_lys %>%
  mutate( snow_spatial_unscaled = snow_spatial * snow_spatial_sd + snow_spatial_mean)


# Ensure your quantile variable has correct labels
preds_lys$linear_quantile_label <- dplyr::recode(preds_lys$linear_quantile,
                                                "-1.13"  = "10%",
                                                "1.04" = "90%")
#And then the actual plot
linear_young_snow<-ggplot(preds_lys, aes(x = snow_spatial_unscaled,
                     y = predicted,
                     colour = linear_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = linear_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_linear,
                      name = "Linear Index Quantile") +
  scale_fill_manual(values = okabe_ito_linear,
                    name = "Linear Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("10 Year Average Winter\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(young_axis_limits)

########################################################################################
#And now predictions for the Snow Anomaly
preds_lya<-data.frame(ggeffects::ggpredict(
             linear_young,
  terms = c("snow_anomaly [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         linear_quantile = group))

#Back transform the snow variable
preds_lya<- preds_lya %>%
  mutate( snow_anomaly_unscaled = snow_anomaly * snow_anomaly_sd + snow_anomaly_mean)

hist(preds_lya$snow_anomaly_unscaled)

# Ensure your quantile variable has correct labels
preds_lya$linear_quantile_label <- dplyr::recode(preds_lya$linear_quantile,
                                                "-1.13"  = "10%",
                                                "1.04" = "90%")
#And then the actual plot
linear_young_anomaly<-ggplot(preds_lya, aes(x = snow_anomaly_unscaled,
                     y = predicted,
                     colour = linear_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = linear_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_linear,
                      name = "Linear Index Quantile") +
  scale_fill_manual(values = okabe_ito_linear,
                    name = "Linear Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("Snow Anomaly\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(young_axis_limits)

########################################################################################
########################################################################################
########################################################################################
#Polygonal model and predictions
polygonal_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_young)


#Make predictions from that model
preds_pys<-data.frame(ggeffects::ggpredict(
             polygonal_young,
  terms = c("mean_snow_spatial [all]",
            paste0("polygonal_index [", round(polygonal_QL, 2), ",", round(polygonal_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         polygonal_quantile = group))

#Back transform the snow variable
preds_pys<- preds_pys %>%
  mutate( snow_spatial_unscaled = snow_spatial * snow_spatial_sd + snow_spatial_mean)

hist(preds_pys$snow_spatial_unscaled)

# Ensure your quantile variable has correct labels
preds_pys$polygonal_quantile_label <- dplyr::recode(preds_pys$polygonal_quantile,
                                                "-0.93"  = "10%",
                                                "1.39" = "90%")
#And then the actual plot
polygonal_young_snow<-ggplot(preds_pys, aes(x = snow_spatial_unscaled,
                     y = predicted,
                     colour = polygonal_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = polygonal_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_polgyonal,
                      name = "Polygonal Index Quantile") +
  scale_fill_manual(values = okabe_ito_polgyonal,
                    name = "Polygonal Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("10 Year Average Winter\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(young_axis_limits)

#And now the anomaly plots
#Polygonal model and predictions


#Make predictions from that model
preds_pya<-data.frame(ggeffects::ggpredict(
             polygonal_young,
  terms = c("snow_anomaly [all]",
            paste0("polygonal_index [", round(polygonal_QL, 2), ",", round(polygonal_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         polygonal_quantile = group))

#Back transform the snow variable
preds_pya<- preds_pya %>%
  mutate( snow_anomaly_unscaled = snow_anomaly * snow_anomaly_sd + snow_anomaly_mean)

hist(preds_pya$snow_anomaly_unscaled)

# Ensure your quantile variable has correct labels
preds_pya$polygonal_quantile_label <- dplyr::recode(preds_pya$polygonal_quantile,
                                                "-0.93"  = "10%",
                                                "1.39" = "90%")
#And then the actual plot
polygonal_young_anomaly<-ggplot(preds_pya, aes(x = snow_anomaly_unscaled,
                     y = predicted,
                     colour = polygonal_quantile_label)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high,
                  fill = polygonal_quantile_label),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = okabe_ito_polgyonal,
                      name = "Polygonal Index Quantile") +
  scale_fill_manual(values = okabe_ito_polgyonal,
                    name = "Polygonal Index Quantile") +
  ylab("Predicted Probability of\nWeekly Occurrence ± 95% CI") +
  xlab("Snow Anomaly\nSnow-Water Equivalent (kg/m²)") +
  theme_classic(base_size = 15)+
  ylim(young_axis_limits)

```


Arrange and export the figures here
```{r}
############################################################################
#Male plots
linear_male_snow
linear_male_anomaly
polygonal_male_snow
polygonal_male_anomaly

#This is greasy, but here I am arranging two arranged plots twice LOL
male_models<-ggpubr::ggarrange(
  
  #Arrange the linear models together with their common legend
  ggpubr::ggarrange(linear_male_snow, linear_male_anomaly,
                  nrow = 2,
                  common.legend = TRUE,
                  legend = "top",
                  labels = c("A", "B"))
                  ,
 #Arrange the polygonal models together with their common legend
 ggpubr::ggarrange(polygonal_male_snow, polygonal_male_anomaly,
                  nrow = 2,
                  common.legend = TRUE,
                  legend = "top",
                  labels = c("C", "D")),

ncol = 2

)

#And write that figure
ggsave(filename = "male_models.jpeg", #name
        plot = male_models,  #plot you want to save
        path = figure_directory, #where its saved
        width = 3000, #how wide
        height = 3000,#how tall
         units = "px", #units in pixels
         bg = 'white') #make sure background is white

####################################################################################
#Female plots
linear_female_snow
linear_female_anomaly
polygonal_female_snow
polygonal_female_anomaly

#This is greasy, but here I am arranging two arranged plots twice LOL
female_models<-ggpubr::ggarrange(
  
  #Arrange the linear models together with their common legend
  ggpubr::ggarrange(linear_female_snow, linear_female_anomaly,
                  nrow = 2,
                  common.legend = TRUE,
                  legend = "top",
                  labels = c("A", "B"))
                  ,
 #Arrange the polygonal models together with their common legend
 ggpubr::ggarrange(polygonal_female_snow, polygonal_female_anomaly,
                  nrow = 2,
                  common.legend = TRUE,
                  legend = "top",
                  labels = c("C", "D")),

ncol = 2

)

#And write that figure
ggsave(filename = "female_models.jpeg", #name
        plot = female_models,  #plot you want to save
        path = figure_directory, #where its saved
        width = 3000, #how wide
        height = 3000,#how tall
         units = "px", #units in pixels
         bg = 'white') #make sure background is white




####################################################################################
#Female plots
linear_young_snow
linear_young_anomaly
polygonal_young_snow
polygonal_young_anomaly

#This is greasy, but here I am arranging two arranged plots twice LOL
young_models<-ggpubr::ggarrange(
  
  #Arrange the linear models together with their common legend
  ggpubr::ggarrange(linear_young_snow, linear_young_anomaly,
                  nrow = 2,
                  common.legend = TRUE,
                  legend = "top",
                  labels = c("A", "B"))
                  ,
 #Arrange the polygonal models together with their common legend
 ggpubr::ggarrange(polygonal_young_snow, polygonal_young_anomaly,
                  nrow = 2,
                  common.legend = TRUE,
                  legend = "top",
                  labels = c("C", "D")),

ncol = 2

)

#And write that figure
ggsave(filename = "young_models.jpeg", #name
        plot = young_models,  #plot you want to save
        path = figure_directory, #where its saved
        width = 3000, #how wide
        height = 3000,#how tall
         units = "px", #units in pixels
         bg = 'white') #make sure background is white


```



#####################################################################################


Now for males
```{r}

#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_males)

#What about a "null" model that doesn't include disturbance?
null_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_males)



#All disturbance
# disturbance_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
#                    disturbance_index+
#                    PC1 + PC2 + #Habitat covariates
#                    predator_index + #Predator index
#                    lat+
#                    mean_snow_spatial*disturbance_index+
#                    snow_anomaly*disturbance_index+
#                    (1|array),
#                  family = "binomial",
#                  data = wtd_dets)
# summary(disturbance_males)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_males)


#Linear
linear_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_males)



#Model selection and cleanup into a nice dataframe!
model_selection_male<- as.data.frame(model.sel(global_males, null_males, linear_males, polygonal_males,
              rank = "AICc"))

model_selection_male<-model_selection_male%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_male)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_male)<-NULL




```


Females
```{r}
#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_females)

#What about a "null" model that doesn't include disturbance?
null_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_females)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_females)


#Linear
linear_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_females)



#Model selection and cleanup into a nice dataframe!
model_selection_females<- as.data.frame(model.sel(global_females, null_females, linear_females, polygonal_females,
              rank = "AICc"))

model_selection_females<-model_selection_females%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_females)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_females)<-NULL




```


Young Deer?
```{r}
#All disturbance
#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_young)

#What about a "null" model that doesn't include disturbance?
null_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_young)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_young)


#Linear
linear_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_young)



#Model selection and cleanup into a nice dataframe!
model_selection_young<- as.data.frame(model.sel(global_young, null_young, linear_young, polygonal_young,
              rank = "AICc"))

model_selection_young<-model_selection_young%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_young)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_young)<-NULL



```


Ok, in the following here is what I want to do. I want to write a loop that performs an out of sample test of our model selection, to identify where the "best" top model is identified
```{r}

#These are the response variables to iterate over
response_variables<-data.frame(response = c("cbind(n_male, (n_weeks - n_male)) ~",
                      "cbind(n_female, (n_weeks - n_female)) ~",
                      "cbind(n_young, (n_weeks - n_young)) ~"),
                      name = c("males", "females", "young"))


#Create a blank list of model selection tables
model_selection_tables<-list()

#How many simulations to do?
n_sims<-100



for(j in 1:nrow(response_variables)){
    
  #Store the current response variable we are working on 
  response <- as.character(response_variables[j, 1])
  #And the name, we will need that for later
  name<- as.character(response_variables[j, 2])
  
  print(paste("Working on models for", name, "at", Sys.time()))
  
  
  #Pre-emptive, but its more efficient to define the model formulas here within this loop, so its only done once per group rather than everytime within the next loop! #Efficiency 101 baby
  formula_global<- as.formula(paste(response,
                      "polygonal_index+linear_index+PC1 + PC2 +predator_index +lat+mean_snow_spatial+snow_anomaly+
                           + (1|array)", sep = ""))
  
  formula_null<- as.formula(paste(response,
                      "PC1 + PC2 +predator_index +lat+mean_snow_spatial+snow_anomaly+
                           + (1|array)", sep = "")) 
  formula_linear<- as.formula(paste(response,
                      "linear_index+
                      PC1 + PC2 +predator_index +lat+
                      mean_snow_spatial*linear_index+
                      snow_anomaly*linear_index+
                           + (1|array)", sep = ""))
  formula_polygonal<- as.formula(paste(response,
                      "polygonal_index+
                      PC1 + PC2 +predator_index +lat+
                      mean_snow_spatial*polygonal_index+
                      snow_anomaly*polygonal_index+
                           + (1|array)", sep = ""))
  
  
  #Now that we have identified our response variable, we need to many
  #many times sample the data, fit each model, run model selection,
  #and identify the top model. 
  for(i in 1:n_sims){
    
    #Sample the dataframe - THIS IS WHERE WE DEFINE THE SAMPLING AMOUNT
    sample<-wtd_dets%>%
      slice_sample(prop = 0.90)
    
    
    #Fit each model!
    fit_global<-glmmTMB::glmmTMB(
                formula = formula_global,
                 family = "binomial",
                 data = sample)
    
    fit_null<-glmmTMB::glmmTMB(
                formula = formula_null,
                 family = "binomial",
                 data = sample)
    
    fit_linear<-glmmTMB::glmmTMB(
                formula = formula_linear,
                 family = "binomial",
                 data = sample)
    
    fit_polygonal<-glmmTMB::glmmTMB(
                formula = formula_polygonal,
                 family = "binomial",
                 data = sample)
    
    #Quick model selection! and cleanup
   model_sel<-as.data.frame(MuMIn::model.sel(fit_global, fit_null, fit_linear, fit_polygonal,
              rank = "AICc"))
   
   model_sel<-model_sel%>%
    # Add model names as a column
  mutate(model = rownames(model_sel)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))
   
   #Get rid of the rownames, they are annoying
   rownames(model_sel)<-NULL
   
   #Add the name, and the simulation number to this dataframe
   model_sel<-model_sel%>%
     mutate(group = name,
            simulation = i)%>%
     relocate(group, simulation)
   
   # Initialize the nested list if this is the first simulation for this group
if (! name %in% names(model_selection_tables)) {
  model_selection_tables[[name]] <- list()
}

# Store the model table for this simulation
model_selection_tables[[name]][[i]] <- model_sel
  }
  
  #Nothing to see here!
  
}

#Ok now lets take those results and strip them out into a single dataframe
all_model_results <- purrr::map_df(
  model_selection_tables,
  ~ purrr::map_df(.x, ~ .x)
)

#And finally, for each group, how often was the top model top?
all_models <- c("fit_linear", "fit_polygonal", "fit_global", "fit_null")


summary_table <-
  all_model_results %>%
  # Only the top model per simulation
  filter(delta == 0) %>%
  group_by(group, model) %>%
  summarise(n_obs = n(), .groups = "drop") %>%
  
  # Expand so every group has all 4 models
  tidyr::complete(
    group,
    model = all_models,
    fill = list(n_obs = 0)
  ) %>%
  
  # Add total sims and proportion
  mutate(
    total = n_sims,
    prop_top = n_obs / total
  )

#And a plot!
# Set factor order
summary_table <- summary_table %>%
  mutate(
    group = factor(group, levels = c("males", "females", "young")),
    model = factor(model,
                   levels = c("fit_linear", "fit_polygonal", "fit_global", "fit_null"),
                   labels = c("Linear", "Polygonal", "Global", "Null"))
  )

ggplot(summary_table, aes(x = model, y = prop_top, fill = model)) +
  geom_col(width = 0.7, alpha = 0.9) +
  facet_wrap(~ group) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     limits = c(0, 1)) +
  labs(
    title = "Proportion of Simulations in Which Each Model Was Top",
    x = "Model",
    y = "Proportion Top Model"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )

```
