The purpose of this file is to do the setup of the response varible for deer? Lets do this clean. 


```{r}
#Clear everything out and start fresh
rm(list=ls())
library(reproducible) #setting up directories for data management
library(readr)        #Needed to read and merge all the csv's in the camera folder
library(dplyr)        #data wranglin'
library(lubridate)    #datetime manipulation
library(tidyr)        #more data manipulation
library(stringr)      #detecting character strings
library(ggplot2)      #data visualization
library(purrr)        #feline-friendly loops 
library(corrplot)     #correlation tests
library(factoextra)   #Principle Components Analysis
library(rgl)


#This just sets up files for inputs (already created, must contain your files), and outputs
input_directory<-reproducible::checkPath(file.path(getwd(), "1. Species Response Variables/outputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "2. Site Variables/outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "2. Site Variables/figures"), create = TRUE)


#read in the files we need
#This is from the species response variable file
wtd_dets<-read.csv(file.path(input_directory, "weekly_proportional_binomial_with_males.csv"))

#These are from the original input folder (fix later to be better)
camop<-read.csv(file.path(getwd(), "1. Species Response Variables/inputs/OSM_camera_deployment.csv"))
covs<-read.csv(file.path(getwd(), "1. Species Response Variables/inputs/OSM_camera_covariates.csv"))

#And the weather data (needs to update)
weather<-read.csv(file.path(getwd(), "1. Species Response Variables/inputs/daymet_weather_data.csv"))

#I am reading this in from the desktop since it is too large to push to github
#Note Aidan you will have to update this
weather_10<-read.csv("C:/Users/16043/Desktop/OSM white tailed deer/daymet_weather_data_10year.csv")





```




First thing I would like to do is construct a Principle Components Analysis of the habitat variables
```{r}
#Just as a side note here, later on in the analyses I remove an observation from the dataset due to an outlier. This does not impact the PCA scores calculated here, since the observation removed is LU13_03 from 2022, but LU13_03 from 2023 still exists. Given that the site is represented in the dataset here, and the habitat variables do not change from year to year, we are good. - Barnas, Feb 4/26


#First thing to do is to extract and rename the natural habitat variables from the dataframe
#Naming conventions are weird from the ABMI source, going to update based on common names

habitat<-covs%>%
  #Using 1000m buffer to match the weather variables 
  filter(buff_dist == 1000)%>%
  #Rename them variables
  dplyr::select(array, site,
                grassland = lc_class110,
                water = lc_class20,
                coniferous = lc_class210,
                broadleaf = lc_class220,
                mixed = lc_class230,
                shrub = lc_class50
                )%>%
  #And scale them all at once
  mutate(across(c(water, broadleaf, grassland, coniferous, mixed, shrub),
              .fns = ~ as.numeric(scale(.x))))


# Run PCA (center = FALSE, scale. = FALSE since theyre already scaled)
habitat_pca <- prcomp(habitat %>%
                        dplyr::select(water, broadleaf, grassland, coniferous, mixed, shrub),
                      center = FALSE, scale. = FALSE)

# Inspect PCA results - the cumulative proportion bascially tells you the variance explained by the cumulative addition of each principle component
summary(habitat_pca)
# Loadings (how variables contribute to PCs), basically as we move from each component how do habitats change?
habitat_pca$rotation

#Save that as a dataframe and export it. Again, this is not affected by the outlier removal later
write.csv(data.frame(habitat_pca$rotation),
          file.path(output_directory, "PCA_results.csv"),
          row.names = TRUE)

#E.g. PC1 broadleaf = 0.55, strong positive contribution, coniferous = -0.69 strong negative contribution
#PC1 represents a gradient between broadleaf and coniferous cover --> high PC1 more broadleaf less coniferous, low PC1 more coniferous, less broadleaf
#PC2 gradient from mixed to shrubby



# PCA scores for each site - these are the values we add back into the dataframe
head(habitat_pca$x)

# Combine PCA scores back with site info (use this to select PCs later)
habitat_pca_df <- bind_cols(
  habitat %>% dplyr::select(array, site),
  as.data.frame(habitat_pca$x)
)

# Quick visualization of the first two PCs
ggplot(habitat_pca_df, aes(x = PC1, y = PC2, color = array)) +
  geom_point(size = 3) +
  theme_minimal() +
  labs(title = "PCA of Habitat Variables",
       x = "PC1 (coniferous to broadleaf)", y = "PC2 (mixed to shrub)")


PCA_plot<-fviz_pca_biplot(habitat_pca,
                axes = c(1, 2),     # use first 2 
                label = "var",         # show variable loadings
                habillage = habitat$array,  # color by array
                addEllipses = TRUE, 
                repel = TRUE,
                geom.ind = "point") +
  ggtitle("3D PCA Biplot of Habitat Variables")

#Save that figure 
ggsave(filename = "PCA_plot.jpeg", #name
        plot = PCA_plot,  #plot you want to save
        path = figure_directory, #where its saved
        width = 2000, #how wide
        height = 2000,#how tall
         units = "px", #units in pixels
         bg = 'white') #make sure background is white


#Ok lets use PC1 and PC2 for now, lets put them into the main dataframe
#Might need to consider PC3s later
wtd_dets<-wtd_dets%>%
  left_join(habitat_pca_df%>%
              dplyr::select(array, site, PC1, PC2),
            by = c("array", "site"))

```

##############################################################################

Alright lets continue to work on our covariate file. I am going to keep things very simple for the time being. Just looking at a few 
```{r}
#I think I want to experiment with a few key ones, along with an overall habitat disturbance score
#Based off Marissas SEM results: 
#Positive: roads, wells (active), wells (inactive), timber harvest
#Negative: Seismic lines, 3d seismic lines, pipelines
colnames(covs)

#Lets pull those out
covs<-covs%>%
  filter(buff_dist == 1000)%>%
  relocate(array, site)%>%
  #remove the natural habitat classifications
  select(-c(lc_class110, lc_class20, lc_class210, lc_class220, lc_class230,lc_class50))%>%
  #keeping a few of these that could be considered disturbance
  rename(rock_and_rubble = lc_class32,
         agriculture = lc_class120,
         exposed = lc_class33,
         developed = lc_class34
         )

#Removing the reclaimed stuff since that should not be considered disturbance (likely little impact  since it looks like its mostly 0's in the dataset - BUT FOR CONSISTENCY I GUESS!)
covs<-covs%>%
  dplyr::select(-c(ris_reclaimed_certified, ris_reclaimed_permanent, ris_reclaimed_temp))

#Also removing features where it is not clear if it would fit into a polygonal or linear category
covs<-covs%>%
  dplyr::select(-c(surrounding_veg, transfer_station))

#Create a sum across all the columns of disturbance, this will represent an overall disturbance score!
covs<-covs%>%
  mutate(disturbance_index = rowSums(across(4:106)))

#I also want to create a broad "linear" and "polygonal" sum variable. This is just a sum of the polygonal and linear sums as I read the definitions in the file.
covs<-covs%>%
  mutate(
    linear_index = airp_runway + canal + conventional_seismic + interchange_ramp + low_impact_seismic + 
      pipeline + ris_airp_runway + ris_road + ris_transmission_line + ris_windrow + rlwy_dbl_track + rlwy_mlt_track + rlwy_sgl_track + rlwy_spur + road_gravel_1l + road_gravel_2l + road_paved_1l + road_paved_2l + road_paved_3l + road_paved_4l + road_paved_5l + road_paved_div + road_paved_undiv_1l  + road_unclassified + road_unimproved + road_unpaved_1l + road_unpaved_2l + road_paved_undiv_2l + road_winter + runway + trail + transmission_line + truck_trail + vegetated_edge_railways + vegetated_edge_roads
    ,  
    
    polygonal_index = borrowpit_dry + borrowpit_wet + borrowpits + camp_industrial + campground + cfo + clearing_unknown + clearing_wellpad_unconfirmed + country_residence + crop + cultivation_abandoned + dugout + facility_other + facility_unknown + fruit_vegetables + golfcourse + greenspace + grvl_sand_pit + harvest_area + harvest_area_white_zone + lagoon + landfill + rock_and_rubble + agriculture + exposed + developed + mill + mines_oilsands + mines_pitlake + misc_oil_gas_facility + oil_gas_plant + peat +  open_pit_mine + recreation + reservoir + residence_clearing + ris_borrowpits + ris_camp_industrial + ris_clearing_unknown + ris_drainage + ris_facility_operations + ris_facility_unknown + ris_mines_oilsands + ris_oilsands_rms + ris_overburden_dump + ris_plant + ris_reclaim_ready + ris_soil_replaced + ris_soil_salvaged + ris_tailing_pond + ris_tank_farm + ris_utilities + ris_waste + rough_pasture + rural_residence + tailing_pond + tame_pasture + urban_industrial + urban_residence + well_aband + well_bitumen + well_cased + well_cleared_not_confirmed + well_cleared_not_drilled + well_gas + well_oil + well_other + well_unknown + sump
  )

#############################################

#Ok now we can work on the specific ones
colnames(covs)

covs<-covs%>%
  #Lets put all the roads together (along with their associated vegetated edges)
  mutate(roads = road_gravel_1l + road_gravel_2l + road_paved_1l + road_paved_2l + road_paved_3l + road_paved_4l + road_paved_5l + road_paved_div + road_paved_undiv_1l + road_paved_undiv_2l + road_unclassified + road_unimproved + road_unpaved_1l + road_unpaved_2l + road_winter + vegetated_edge_roads,
         
  #different well types, along with a generic "well"
  #Not including "well_unknown" since that states the status is unknown
  well_gas = well_gas,
  well_oil = well_oil,
  well_bitumen = well_bitumen,
  wells = well_gas + well_oil + well_bitumen + well_other,
  inactive_well = well_aband,
  
  seismic = conventional_seismic,
  seismic_3D = low_impact_seismic,
  harvest = harvest_area,
  pipeline = pipeline)%>%
  #And then just keep the ones we care about
  dplyr::select(array, site, disturbance_index, linear_index, polygonal_index, roads, wells, well_gas, well_oil, well_bitumen, inactive_well, seismic, seismic_3D, harvest, pipeline)

#Some very quick and dirty exploration
hist(covs$disturbance_index)
hist(covs$well_gas)
hist(covs$well_oil)
hist(covs$well_bitumen)
hist(covs$wells)
hist(covs$roads)
hist(covs$inactive_well)
hist(covs$seismic)
hist(covs$seismic_3D)  
hist(covs$harvest)
hist(covs$pipeline)       

#Recognize that we still ahve all our sites!
n_distinct(covs$site)



#Lets link that into the detection file
wtd_dets<-merge(wtd_dets, covs, by = c("site", "array"))


```

Weather data. Ok there are a few different things that need doing in this block. Going to keep it all simple and just work with snow for now. But I need to build:
1) RAW: For each site - the preceeding years mean snow data
2) AVERAGE TEMPORAL: for each site, the average mean snow within an array in the preceeding year
3) AVERAGE SPATIAL: for each site, the average mean snow for that site across the last 10 years
4) ANOMALY: for each site the difference between the current data - temporal and - spatial.
```{r}
#So first things first, lets keep life simple and just look at snow
weather_10<-weather_10%>%
  dplyr::select(c(site, latitude, longitude, year, day, snow = snow_water_equiv_kgM2), temp = min_temp_C)

#I would like a couple graphs here showing the time series for each variable
#First lets get the dates read in nicely
weather_10 <- weather_10 %>%
  mutate(date = ymd(paste0(year, "-01-01")) + days(day - 1))

weather_summary<-weather_10%>%
  group_by(date)%>%
  summarise(mean_snow = mean(snow),
            sd_snow = sd(snow))

#Snow Plot
ggplot(weather_summary, aes(x = date)) +
  geom_ribbon(
    aes(ymin = mean_snow - sd_snow,
        ymax = mean_snow + sd_snow),
    fill = "grey", alpha = 0.6
  ) +
  geom_line(aes(y = mean_snow), color = "black", size = 1) +
  ylab("Mean Snow Water Equivalent (kg)") +
  xlab("Date") +
  ggtitle("Daily Snow")+
  theme_classic()

######################################################################################
#Ok, lets get these measures into the dataframe. Now, we want to think about lag times and whatnot. So lets extract 

#First lets define the winter months (using guidance from Dickie et al which reports snow periods as Sep 1 to March 31) (Potentially rethink this since we use September data for fawns)
winter_months<-c(9, 10, 11, 12, 1, 2, 3) 


#First just keep data in those months
weather_10<-weather_10%>%
  filter(month(date) %in% winter_months)
  

#Ok ok. Now remember, the weather dataframe has data for all sites for all years. But we don't need that 
weather_summary <- weather_10 %>%
  # Extract month for filtering
  mutate(month = month(date),
         # Define "winter_year" = the year that the winter *starts*
         winter_year = if_else(month >= 9, year(date), year(date) - 1)) %>%
  
  # Group by site and winter season
  group_by(site, winter_year) %>%
  
  # Summarize mean and sd for each variable
  summarise(
    mean_snow  = mean(snow, na.rm = TRUE),
    mean_min_temp   = mean(temp, na.rm = TRUE),
    n_days    = n(),  # optional — good to check how many days were available
    .groups = "drop"
  )

#Ok so at this point I have the winter_year data for each site. Now I need to manipulate and decompose that data into what I need.




#First, lets get the raw data into the wtd_dets dataframe
#Identify the winter year to use
wtd_dets<-wtd_dets%>%
  mutate(winter_year = year - 1)

############################################
#ADDING IN A LIL CHUNK FOR LAG YEARS########
#I think the simplest way will be to create extra dataframes and add them in sort of manually

lag2<-wtd_dets%>%
  #two winters before
  mutate(winter_year = year - 2)%>%
  dplyr::select(c(site, winter_year, year))%>%
  left_join(weather_summary%>%
              dplyr::select(site, winter_year, mean_snow))%>%
  rename(winter_year_lag2 = winter_year,
         mean_snow_lag2 = mean_snow)%>%
  #Dumb, but I need to add back in the original winter year for re-joining later
  mutate(winter_year = year - 1)

lag3<-wtd_dets%>%
  #two winters before
  mutate(winter_year = year - 3)%>%
  dplyr::select(c(site, winter_year, year))%>%
  left_join(weather_summary%>%
              dplyr::select(site, winter_year, mean_snow))%>%
  rename(winter_year_lag3 = winter_year,
         mean_snow_lag3 = mean_snow)%>%
  #Dumb, but I need to add back in the original winter year for re-joining later
  mutate(winter_year = year - 1)

############################################
wtd_dets<-wtd_dets%>%
  left_join(weather_summary%>%
  dplyr::select(site, winter_year, mean_snow, mean_min_temp), 
  by = c("site", "winter_year"))


#And then add in the lagged data
wtd_dets<-wtd_dets%>%
  left_join(lag2)%>%
  left_join(lag3)


#make sure that worked- no NAs!
summary(wtd_dets$mean_snow)
hist(wtd_dets$mean_snow)
summary(wtd_dets$mean_min_temp)
hist(wtd_dets$mean_min_temp)

###########################################
#Ok now the temporal component. For each site/year what was the average winter snow across its respective array.
#To do this I will need to add array back into the weather data. My bad. I can get that from camop

weather_summary<-weather_summary%>%
  left_join(camop%>%
  dplyr::select(site, array),
  by = "site")%>%
  relocate(array, site)

#So this calculation is slightly different! Grouping by year to get an overall average!
weather_summary%>%
  group_by(winter_year)%>%
  summarise(mean_snow_temporal = mean(mean_snow),
            mean_min_temp_temporal = mean(mean_min_temp))


#join it all up
wtd_dets<-wtd_dets%>%
  left_join(
weather_summary%>%
  group_by(winter_year)%>%
  summarise(mean_snow_temporal = mean(mean_snow),
            mean_min_temp_temporal = mean(mean_min_temp)),
by = c("winter_year"))

###################################################################
#Average spatial - so now we are going to take a single site, and find the average over the last 10 years
head(weather_summary)

#This is slightly tricky though, because I need to find a date range for each site. I've got a rather clever idea. Lets take the winter year, and then calculate a min year 10 years before that
site_years<-wtd_dets%>%
  mutate(min_year = winter_year - 10)%>%
  dplyr::select(site, winter_year, min_year)

site_10_year_snow<-site_years %>%
  left_join(weather_summary, by = "site") %>%
  #determining which years to keep based on the winter years in both frames
   #keep weather summary winter year if its greater than the minimum for its site
   filter(winter_year.y >= min_year, 
         #keep weather summary winter year if its less than the max for its site  
         winter_year.y <= winter_year.x) %>%
  group_by(site, winter_year = winter_year.x) %>%
  summarise(mean_snow_spatial = mean(mean_snow, na.rm = TRUE),
            mean_min_temp_spatial = mean(mean_min_temp, na.rm = TRUE), .groups = "drop")

#And now we can join that in e z p z
wtd_dets<-wtd_dets%>%
  left_join(site_10_year_snow, by = c("site", "winter_year"))

############################################################################
#Winter anomaly!

#This one should be easier I think to calculate as I am pretty sure it is just the difference between the raw and the time minus space
wtd_dets<-wtd_dets%>%
  mutate(snow_anomaly = mean_snow - mean_snow_spatial,
         temp_anomaly = mean_min_temp - mean_min_temp_spatial)

hist(wtd_dets$snow_anomaly)
hist(wtd_dets$temp_anomaly)




```

############################################################################################################
Alright at this point, everyone is "in" the dataframe. So now before I do any calculations below I need to remove an important outlier that I discovered. LU13_03 in 2022 has a large outlier for predator detections, so unfortunately I am going to remove that whole observation here. This way it doesn't impact calculations downstream.
```{r}

wtd_dets <- wtd_dets %>%
  filter(!(site == "LU13_03" & year == 2022))

#Sanity check, redundant but important
wtd_dets%>%
  filter(site == "LU13_03" & year == 2022)


```

Just doing a quick correlation test here between snow and temperature. Using this as a justification for just examining snow and not temperature in the manuscript.
```{r}
#This is the key thing I reference in the paper
cor.test(wtd_dets$mean_snow_spatial, wtd_dets$mean_min_temp_spatial)



#Give me some correlations between winter and temperature?
cor.test(wtd_dets$mean_min_temp, wtd_dets$mean_snow)

cor.test(wtd_dets$snow_anomaly, wtd_dets$temp_anomaly)



```


Predator detections?
```{r}
hist(wtd_dets$predator_detections)
#Lets make this into a rate controlling for effort
wtd_dets<-wtd_dets%>%
  mutate(predator_index = predator_detections/n_weeks)

hist(wtd_dets$predator_index)

#Lazy but add the site coordinates into the final file
wtd_dets<-wtd_dets%>%
  left_join(camop%>%dplyr::select(site, lat, long), by = "site")%>%
  relocate(year, array, site, lat, long)

```

So now we have our "final" dataframe, lets put together a figure of the data
```{r}

#Variables to plot
vars_of_interest <- c(
  "predator_index", "PC1", "PC2", "linear_index", 
  "polygonal_index", "mean_snow", "snow_anomaly", "lat"
)

#Custom labels
x_labels <- c(
  predator_index   = "Predator detection index",
  PC1              = "Habitat gradient (PC1)",
  PC2              = "Habitat gradient (PC2)",
  linear_index     = "Linear disturbance index",
  polygonal_index  = "Polygonal disturbance index",
  mean_snow        = "Long term average snow\nSnow-Water Equivalent (kg/m²)",
  snow_anomaly     = "Snow anomaly\nsnow-water equivalent (kg/m²)",
  lat              = "Latitude (°N)"
)

#Loop to make some plots
plot_list <- lapply(vars_of_interest, function(v) {
  
  ggplot(wtd_dets, aes(x = .data[[v]])) +
    geom_histogram(fill = "grey", color = "black", bins = 20) +
    labs(
      x = x_labels[[v]],
      y = "Frequency"
    ) +
    theme_classic() +
    theme(axis.title = element_text(size = 10))
})

#And then arrange them all
variables_plot <- ggpubr::ggarrange(
  plotlist = plot_list,
  ncol = 3, nrow = 3,
  labels = "AUTO"
)


#Export baby!
ggsave(filename = "variables_plot.jpeg", #name
        plot = variables_plot,  #plot you want to save
        path = figure_directory, #where its saved
        width =3000, #how wide
        height = 2000,#how tall
         units = "px", #units in pixels
         bg = 'white') #make sure background is white


```


Write that final dataframe
```{r}
#Lets write that file and move on to modeling!
write.csv(wtd_dets, file.path(output_directory, "final_df.csv"),
          row.names = FALSE)

```
