The purpose of this file is to do the model selection for the manuscript (write better later)

1. main model selection table for manuscript
2. out of sample test for models


```{r}
#Clear everything out and start fresh
rm(list=ls())
library(reproducible) #setting up directories for data management
library(readr)        #Needed to read and merge all the csv's in the camera folder
library(dplyr)        #data wranglin'
library(lubridate)    #datetime manipulation
library(tidyr)        #more data manipulation
library(stringr)      #detecting character strings
library(ggplot2)      #data visualization
library(purrr)        #feline-friendly loops 
library(corrplot)     #correlation tests
library(MuMIn)
library(ggeffects)


#This just sets up files for inputs (already created, must contain your files), and outputs
input_directory<-reproducible::checkPath(file.path(getwd(), "2. Site Variables/outputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "3. Model Selection/outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "3. Model Selection/figures"), create = TRUE)


#read in the files we need
#This is from the species response variable file
wtd_dets<-read.csv(file.path(input_directory, "final_df.csv"))



```

Lets prepare the data for modeling.
```{r}

#PC variables are already scaled, but lets scale the remaining variables
wtd_dets<-wtd_dets%>%
  mutate(across(c(disturbance_index, polygonal_index, linear_index, mean_snow_spatial, snow_anomaly, predator_index, lat), .fns = ~ as.numeric(scale(.x))))

         
#Lets do a quick check of correlations
vars <- wtd_dets %>%
  dplyr::select(disturbance_index, polygonal_index, linear_index, mean_snow_spatial, snow_anomaly, predator_index, lat, PC1, PC2)

# Compute correlation matrix (use complete.obs to ignore NAs)
cor_matrix <- cor(vars, use = "complete.obs")


#And now lets identify the actual correlation values to report in the manuscrpt
cor_df <- as.data.frame(cor_matrix) %>%
  tibble::rownames_to_column(var = "var1") %>%
  pivot_longer(
    cols = -var1,
    names_to = "var2",
    values_to = "correlation"
  )%>%
  # Remove self-correlations and duplicate pairs easy peasy (or is it peazy?)
  filter(var1 < var2)%>%
   # Keep only strong correlations
  filter(abs(correlation) >= 0.5)%>%
  arrange(desc(abs(correlation)))

#Yeah so there is a strong correlation between polygonal features and total disturbance, this makes sense right, since if you think about it, the disturbance measure is the sum of all polygonal and linear features. And since linear features are such a small footprint, the majority of the total disturbance is just polygonal. So we wont model total disturbance, and make sure we talk about that in the manuscript

#What is the actual value to report though
cor.test(wtd_dets$disturbance_index, wtd_dets$polygonal_index)


# Print the correlation matrix now without disturbance index
vars <- wtd_dets %>%
  dplyr::select(#disturbance_index, 
                polygonal_index, linear_index, mean_snow_spatial, snow_anomaly, predator_index, lat, PC1, PC2)
cor_matrix <- cor(vars, use = "complete.obs")
print(round(cor_matrix, 2))

#Produce the plot and save it
png(file.path(figure_directory, "correlation_plot.png"))
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, addCoef.col = "black")  
dev.off()


#Write that dataframe with the scaled variables to the outputs folder (this will be used in the next step for modeling)
write.csv(wtd_dets, file.path(output_directory, "final_df.csv"), row.names = FALSE)

```


Doing some quick summary stuff for camera operation here. No longer including the model seleciton stuff but I am keeping it for now in case I return to it
```{r}
n_distinct(wtd_dets$site)

#how many total survey weeks? That should just be the sum of n_weeks across the frame right?
wtd_dets%>%
  summarise(total_weeks = sum(n_weeks),
            mean_weeks = mean(n_weeks),
            sd_weeks = sd(n_weeks),
            min_weeks = min(n_weeks),
            max_weeks = max(n_weeks))




#Information on site occupancy
 wtd_dets %>%
  group_by(site) %>%
  summarise(
    male_present   = any(n_male   > 0, na.rm = TRUE),
    female_present = any(n_female > 0, na.rm = TRUE),
    young_present  = any(n_young  > 0, na.rm = TRUE),
    .groups = "drop"
  )%>%
   summarise(
    n_sites = n(),
    males   = sum(male_present),
    females = sum(female_present),
    young   = sum(young_present)
  )%>%
  mutate(
    males_prop   = males   / n_sites,
    females_prop = females / n_sites,
    young_prop   = young   / n_sites
  )



```





Now for males
```{r}

#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_males)

#What about a "null" model that doesn't include disturbance?
null_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_males)



#All disturbance
# disturbance_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
#                    disturbance_index+
#                    PC1 + PC2 + #Habitat covariates
#                    predator_index + #Predator index
#                    lat+
#                    mean_snow_spatial*disturbance_index+
#                    snow_anomaly*disturbance_index+
#                    (1|array),
#                  family = "binomial",
#                  data = wtd_dets)
# summary(disturbance_males)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_males)


#Linear
linear_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                     mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_males)



#Model selection and cleanup into a nice dataframe!
model_selection_male<- as.data.frame(model.sel(global_males, null_males, linear_males, polygonal_males,
              rank = "AICc"))

model_selection_male<-model_selection_male%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_male)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_male)<-NULL




```


Females
```{r}
#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_females)

#What about a "null" model that doesn't include disturbance?
null_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_females)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_females)


#Linear
linear_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_females)



#Model selection and cleanup into a nice dataframe!
model_selection_females<- as.data.frame(model.sel(global_females, null_females, linear_females, polygonal_females,
              rank = "AICc"))

model_selection_females<-model_selection_females%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_females)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_females)<-NULL




```


Young Deer?
```{r}
#All disturbance
#Global model - note that overall disturbance and polygonal disturbance are highly correlated....
global_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                     polygonal_index+
                     linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(global_young)

#What about a "null" model that doesn't include disturbance?
null_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   #  polygonal_index+
                  #   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   lat+
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(null_young)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_young)


#Linear
linear_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                     lat+
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_young)



#Model selection and cleanup into a nice dataframe!
model_selection_young<- as.data.frame(model.sel(global_young, null_young, linear_young, polygonal_young,
              rank = "AICc"))

model_selection_young<-model_selection_young%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_young)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_young)<-NULL



```

Nicely format those model seleciton tables, this is a key result!
```{r}
model_selection_male
model_selection_females
model_selection_young


#--- Helper function to clean each table ---
clean_table <- function(df, group_name) {
  df %>%
    mutate(
      group = group_name,

      # Extract just the model part before the underscore
      model = str_extract(model, "^[^_]+"),

      # Convert to nice names
      model = factor(model,
                     levels = c("linear", "polygonal", "global", "null"),
                     labels = c("Linear", "Polygonal", "Global", "Null"))
    )
}

# Clean all three
male_clean    <- clean_table(model_selection_male,    "Males")
female_clean  <- clean_table(model_selection_females, "Lone Females")
young_clean   <- clean_table(model_selection_young,   "Females With Young")

combined_models <- bind_rows(male_clean, female_clean, young_clean)%>%
  relocate(group)%>%
  mutate(logLik = round(logLik, digits = 2),
         AICc = round(AICc, digits = 2),
         delta = round(delta, digits = 2))




library(flextable)
library(officer)

# Convert to a nice Word table
ft <- flextable::flextable(combined_models)

# Optimize the widths and heights of columns. Nice to have!
ft <- flextable::autofit(ft)

#And then turn that into an actual word document
doc <- officer::read_docx()
doc <- flextable::body_add_flextable(doc, ft)

#Now actually write the file
print(doc, target = file.path(output_directory, "model_selection_results.docx"))

```




Ok, in the following here is what I want to do. I want to write a loop that performs an out of sample test of our model selection, to identify where the "best" top model is identified
```{r}

#These are the response variables to iterate over
response_variables<-data.frame(response = c("cbind(n_male, (n_weeks - n_male)) ~",
                      "cbind(n_female, (n_weeks - n_female)) ~",
                      "cbind(n_young, (n_weeks - n_young)) ~"),
                      name = c("males", "females", "young"))


#Create a blank list of model selection tables
model_selection_tables<-list()

#How many simulations to do?
n_sims<-100


#Random seed
set.seed(420)
for(j in 1:nrow(response_variables)){
    
  #Store the current response variable we are working on 
  response <- as.character(response_variables[j, 1])
  #And the name, we will need that for later
  name<- as.character(response_variables[j, 2])
  
  print(paste("Working on models for", name, "at", Sys.time()))
  
  
  #Pre-emptive, but its more efficient to define the model formulas here within this loop, so its only done once per group rather than everytime within the next loop! #Efficiency 101 baby
  formula_global<- as.formula(paste(response,
                      "polygonal_index+linear_index+PC1 + PC2 +predator_index +lat+mean_snow_spatial+snow_anomaly+
                           + (1|array)", sep = ""))
  
  formula_null<- as.formula(paste(response,
                      "PC1 + PC2 +predator_index +lat+mean_snow_spatial+snow_anomaly+
                           + (1|array)", sep = "")) 
  formula_linear<- as.formula(paste(response,
                      "linear_index+
                      PC1 + PC2 +predator_index +lat+
                      mean_snow_spatial*linear_index+
                      snow_anomaly*linear_index+
                           + (1|array)", sep = ""))
  formula_polygonal<- as.formula(paste(response,
                      "polygonal_index+
                      PC1 + PC2 +predator_index +lat+
                      mean_snow_spatial*polygonal_index+
                      snow_anomaly*polygonal_index+
                           + (1|array)", sep = ""))
  
  
  #Now that we have identified our response variable, we need to many
  #many times sample the data, fit each model, run model selection,
  #and identify the top model. 
  for(i in 1:n_sims){
    
    #Sample the dataframe - THIS IS WHERE WE DEFINE THE SAMPLING AMOUNT
    sample<-wtd_dets%>%
      slice_sample(prop = 0.90)
    
    
    #Fit each model!
    fit_global<-glmmTMB::glmmTMB(
                formula = formula_global,
                 family = "binomial",
                 data = sample)
    
    fit_null<-glmmTMB::glmmTMB(
                formula = formula_null,
                 family = "binomial",
                 data = sample)
    
    fit_linear<-glmmTMB::glmmTMB(
                formula = formula_linear,
                 family = "binomial",
                 data = sample)
    
    fit_polygonal<-glmmTMB::glmmTMB(
                formula = formula_polygonal,
                 family = "binomial",
                 data = sample)
    
    #Quick model selection! and cleanup
   model_sel<-as.data.frame(MuMIn::model.sel(fit_global, fit_null, fit_linear, fit_polygonal,
              rank = "AICc"))
   
   model_sel<-model_sel%>%
    # Add model names as a column
  mutate(model = rownames(model_sel)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))
   
   #Get rid of the rownames, they are annoying
   rownames(model_sel)<-NULL
   
   #Add the name, and the simulation number to this dataframe
   model_sel<-model_sel%>%
     mutate(group = name,
            simulation = i)%>%
     relocate(group, simulation)
   
   # Initialize the nested list if this is the first simulation for this group
if (! name %in% names(model_selection_tables)) {
  model_selection_tables[[name]] <- list()
}

# Store the model table for this simulation
model_selection_tables[[name]][[i]] <- model_sel
  }
  
  #Nothing to see here!
  
}
```


And here we are formatting the final figure and whatnot
```{r}
#Ok now lets take those results and strip them out into a single dataframe
all_model_results <- purrr::map_df(
  model_selection_tables,
  ~ purrr::map_df(.x, ~ .x)
)

#And finally, for each group, how often was the top model top?
all_models <- c("fit_linear", "fit_polygonal", "fit_global", "fit_null")


summary_table <-
  all_model_results %>%
  # Only the top model per simulation
  filter(delta == 0) %>%
  group_by(group, model) %>%
  summarise(n_obs = n(), .groups = "drop") %>%
  
  # Expand so every group has all 4 models
  tidyr::complete(
    group,
    model = all_models,
    fill = list(n_obs = 0)
  ) %>%
  
  # Add total sims and proportion
  mutate(
    total = n_sims,
    prop_top = n_obs / total
  )

#And a plot!
# Set factor order
summary_table <- summary_table %>%
  mutate(
    group = factor(group,
                   levels = c("males", "females", "young"),
                   labels = c("Males", "Lone Females", "Females With Young")),
    model = factor(model,
                   levels = c("fit_linear", "fit_polygonal", "fit_global", "fit_null"),
                   labels = c("Linear", "Polygonal", "Global", "Null")))

#Define the colour blind paletee 
okabe_ito <- c(
  "Linear"     = "#0072B2",
  "Polygonal"  = "#E69F00",
  "Global"     = "#009E73",
  "Null"       = "#D55E00"
)

#Save that figure
model_selection_validation_plot<-ggplot(summary_table, aes(x = model, y = prop_top, fill = model)) +
  geom_col(width = 0.7, alpha = 0.9) +
  
  # Add bar labels
  geom_text(aes(label = round(prop_top, 2)),
            vjust = -0.3,            # position above bar
            size = 4,                # text size
            fontface = "bold") +     # optional
  
  facet_wrap(~ group) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     limits = c(0, 1)) +
  scale_fill_manual(values = okabe_ito) +
  labs(
    title = "Proportion of Samples in Which Each Model Was Top",
    x = "Model",
    y = "Proportion Top Model"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )


#Save that figure
#Save that figure 
ggsave(filename = "model_selection_validation_plot.jpeg", #name
        plot = model_selection_validation_plot,  #plot you want to save
        path = figure_directory, #where its saved
        width = 3000, #how wide
        height = 2000,#how tall
         units = "px", #units in pixels
         bg = 'white') #make sure background is white
```
