The purpose of this file is to do the setup of the response varible for deer? Lets do this clean. 


```{r}
#Clear everything out and start fresh
rm(list=ls())
library(reproducible) #setting up directories for data management
library(readr)        #Needed to read and merge all the csv's in the camera folder
library(dplyr)        #data wranglin'
library(lubridate)    #datetime manipulation
library(tidyr)        #more data manipulation
library(stringr)      #detecting character strings
library(ggplot2)      #data visualization
library(purrr)        #feline-friendly loops 
library(corrplot)     #correlation tests
library(MuMIn)
library(ggeffects)


#This just sets up files for inputs (already created, must contain your files), and outputs
input_directory<-reproducible::checkPath(file.path(getwd(), "2. Site Variables/outputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "3. Models/outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "3. Models/figures"), create = TRUE)


#read in the files we need
#This is from the species response variable file
wtd_dets<-read.csv(file.path(input_directory, "final_df.csv"))



```

Lets prepare the data for modeling.
```{r}
#First deal with wolves, need to morph this into a response variable that makes sense

#wtd_dets<-wtd_dets%>%
#  mutate(wolf_index = wolf_present/n_weeks)

#PC variables are already scaled, but lets scale the remaining variables
wtd_dets<-wtd_dets%>%
  mutate(across(c(disturbance_index, polygonal_index, linear_index, roads, wells,
                  well_gas, 
                  well_oil, well_bitumen, inactive_well, seismic, seismic_3D, harvest, pipeline, mean_snow, mean_min_temp, mean_snow_spatial, mean_min_temp_spatial, snow_anomaly, temp_anomaly, predator_index), .fns = ~ as.numeric(scale(.x))))

         
#Lets do a quick check of correlations
vars <- wtd_dets %>%
  dplyr::select(disturbance_index, roads, wells, well_gas, well_oil, well_bitumen, inactive_well, seismic, seismic_3D, harvest, pipeline, mean_snow, mean_min_temp, mean_snow_spatial, mean_min_temp_spatial, snow_anomaly, temp_anomaly, predator_index)

# Compute correlation matrix (use complete.obs to ignore NAs)
cor_matrix <- cor(vars, use = "complete.obs")

# Print the correlation matrix
print(round(cor_matrix, 2))

# Visualize it
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, addCoef.col = "black")       

cor_df <- as.data.frame(cor_matrix) %>%
  tibble::rownames_to_column(var = "var1") %>%
  pivot_longer(
    cols = -var1,
    names_to = "var2",
    values_to = "correlation"
  )%>%
  # Remove self-correlations and duplicate pairs
  filter(var1 < var2)%>%
   # Keep only strong correlations
  filter(abs(correlation) >= 0.6)%>%
  arrange(desc(abs(correlation)))

```


Ok now for each group I want to fit the three models and see what results pop out. Lets deal with all deer first
```{r}

#All disturbance
disturbance_all<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_all)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_all<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_all)


#Linear
linear_all<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_all)



#Model selection and cleanup into a nice dataframe!
model_selection_all<- as.data.frame(model.sel(disturbance_all, linear_all, polygonal_all,
              rank = "AICc"))

model_selection_all<-model_selection_all%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_all)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_all)<-NULL


#Model predictions for the polygonal and the linear
polygon_QL<-quantile(wtd_dets$polygonal_index, 0.10, na.rm = TRUE)
polygon_QH<-quantile(wtd_dets$polygonal_index, 0.90, na.rm = TRUE)

linear_QL<-quantile(wtd_dets$linear_index, 0.10, na.rm = TRUE)
linear_QH<-quantile(wtd_dets$linear_index, 0.90, na.rm = TRUE)

disturbance_QL<-quantile(wtd_dets$disturbance_index, 0.10, na.rm = TRUE)
disturbance_QH<-quantile(wtd_dets$disturbance_index, 0.90, na.rm = TRUE)

#Predictions for each disturbance type and snow interaction
disturbance_spatial_all<-ggeffects::ggpredict(
             disturbance_all,
  terms = c("mean_snow_spatial [all]",
            paste0("disturbance_index [", round(disturbance_QL, 2), ",", round(disturbance_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         disturbance_quantile = group)


polygon_spatial_all<-ggeffects::ggpredict(
             polygonal_all,
  terms = c("mean_snow_spatial [all]",
            paste0("polygonal_index [", round(polygon_QL, 2), ",", round(polygon_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         polygonal_quantile = group)

linear_spatial_all<-ggeffects::ggpredict(
             linear_all,
  terms = c("mean_snow_spatial [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         linear_quantile = group)

m1<-ggplot(disturbance_spatial_all, aes(x = snow_spatial, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("All deer ~ disturbance:mean snow")


m2<-ggplot(polygon_spatial_all, aes(x = snow_spatial, y = predicted, colour = polygonal_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = polygonal_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("All deer ~ polygonal:mean snow")

m3<-ggplot(linear_spatial_all, aes(x = snow_spatial, y = predicted, colour = linear_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = linear_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("All deer ~ linear:mean snow")

#Snow anomaly

disturbance_anomaly_all<-ggeffects::ggpredict(
             disturbance_all,
  terms = c("snow_anomaly [all]",
            paste0("disturbance_index [", round(disturbance_QL, 2), ",", round(disturbance_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         disturbance_quantile = group)


polygon_anomaly_all<-ggeffects::ggpredict(
             polygonal_all,
  terms = c("snow_anomaly [all]",
            paste0("polygonal_index [", round(polygon_QL, 2), ",", round(polygon_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         polygonal_quantile = group)

linear_anomaly_all<-ggeffects::ggpredict(
             linear_all,
  terms = c("snow_anomaly [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         linear_quantile = group)


m4<-ggplot(disturbance_anomaly_all, aes(x = snow_anomaly, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("All deer ~ disturbance:snow anomaly")

m5<-ggplot(polygon_anomaly_all, aes(x = snow_anomaly, y = predicted, colour = polygonal_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = polygonal_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("All deer ~ polygonal:snow anomaly")

m6<-ggplot(linear_anomaly_all, aes(x = snow_anomaly, y = predicted, colour = linear_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = linear_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("All deer ~ linear:snow anomaly")



#Put them all together
ggpubr::ggarrange(m1, m2, m3, m4, m5, m6, nrow = 2, ncol = 3)


```



Now for males
```{r}

#All disturbance
disturbance_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_males)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_males)


#Linear
linear_males<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_males)



#Model selection and cleanup into a nice dataframe!
model_selection_male<- as.data.frame(model.sel(disturbance_males, linear_males, polygonal_males,
              rank = "AICc"))

model_selection_male<-model_selection_male%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_male)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_male)<-NULL


#Model predictions for the polygonal and the linear
polygon_QL<-quantile(wtd_dets$polygonal_index, 0.10, na.rm = TRUE)
polygon_QH<-quantile(wtd_dets$polygonal_index, 0.90, na.rm = TRUE)

linear_QL<-quantile(wtd_dets$linear_index, 0.10, na.rm = TRUE)
linear_QH<-quantile(wtd_dets$linear_index, 0.90, na.rm = TRUE)

disturbance_QL<-quantile(wtd_dets$disturbance_index, 0.10, na.rm = TRUE)
disturbance_QH<-quantile(wtd_dets$disturbance_index, 0.90, na.rm = TRUE)

#Predictions for each disturbance type and snow interaction
disturbance_spatial_males<-ggeffects::ggpredict(
             disturbance_males,
  terms = c("mean_snow_spatial [all]",
            paste0("disturbance_index [", round(disturbance_QL, 2), ",", round(disturbance_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         disturbance_quantile = group)


polygon_spatial_males<-ggeffects::ggpredict(
             polygonal_males,
  terms = c("mean_snow_spatial [all]",
            paste0("polygonal_index [", round(polygon_QL, 2), ",", round(polygon_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         polygonal_quantile = group)

linear_spatial_males<-ggeffects::ggpredict(
             linear_males,
  terms = c("mean_snow_spatial [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         linear_quantile = group)

m1<-ggplot(disturbance_spatial_males, aes(x = snow_spatial, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("Male deer ~ disturbance:mean snow")


m2<-ggplot(polygon_spatial_males, aes(x = snow_spatial, y = predicted, colour = polygonal_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = polygonal_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("Male deer ~ polygonal:mean snow")

m3<-ggplot(linear_spatial_males, aes(x = snow_spatial, y = predicted, colour = linear_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = linear_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("Male deer ~ linear:mean snow")

#Snow anomaly

disturbance_anomaly_males<-ggeffects::ggpredict(
             disturbance_males,
  terms = c("snow_anomaly [all]",
            paste0("disturbance_index [", round(disturbance_QL, 2), ",", round(disturbance_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         disturbance_quantile = group)


polygon_anomaly_males<-ggeffects::ggpredict(
             polygonal_males,
  terms = c("snow_anomaly [all]",
            paste0("polygonal_index [", round(polygon_QL, 2), ",", round(polygon_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         polygonal_quantile = group)

linear_anomaly_males<-ggeffects::ggpredict(
             linear_males,
  terms = c("snow_anomaly [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         linear_quantile = group)


m4<-ggplot(disturbance_anomaly_males, aes(x = snow_anomaly, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("MAle deer ~ disturbance:snow anomaly")

m5<-ggplot(polygon_anomaly_males, aes(x = snow_anomaly, y = predicted, colour = polygonal_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = polygonal_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("Male deer ~ polygonal:snow anomaly")

m6<-ggplot(linear_anomaly_males, aes(x = snow_anomaly, y = predicted, colour = linear_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = linear_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("Male deer ~ linear:snow anomaly")



#Put them all together
ggpubr::ggarrange(m1, m2, m3, m4, m5, m6, nrow = 2, ncol = 3)


```


Females
```{r}
#All disturbance
disturbance_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_females)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_females)


#Linear
linear_females<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_females)



#Model selection and cleanup into a nice dataframe!
model_selection_females<- as.data.frame(model.sel(disturbance_females, linear_females, polygonal_females,
              rank = "AICc"))

model_selection_females<-model_selection_females%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_females)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_females)<-NULL


#Model predictions for the polygonal and the linear
polygon_QL<-quantile(wtd_dets$polygonal_index, 0.10, na.rm = TRUE)
polygon_QH<-quantile(wtd_dets$polygonal_index, 0.90, na.rm = TRUE)

linear_QL<-quantile(wtd_dets$linear_index, 0.10, na.rm = TRUE)
linear_QH<-quantile(wtd_dets$linear_index, 0.90, na.rm = TRUE)

disturbance_QL<-quantile(wtd_dets$disturbance_index, 0.10, na.rm = TRUE)
disturbance_QH<-quantile(wtd_dets$disturbance_index, 0.90, na.rm = TRUE)

#Predictions for each disturbance type and snow interaction
disturbance_spatial_females<-ggeffects::ggpredict(
             disturbance_females,
  terms = c("mean_snow_spatial [all]",
            paste0("disturbance_index [", round(disturbance_QL, 2), ",", round(disturbance_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         disturbance_quantile = group)


polygon_spatial_females<-ggeffects::ggpredict(
             polygonal_females,
  terms = c("mean_snow_spatial [all]",
            paste0("polygonal_index [", round(polygon_QL, 2), ",", round(polygon_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         polygonal_quantile = group)

linear_spatial_females<-ggeffects::ggpredict(
             linear_females,
  terms = c("mean_snow_spatial [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         linear_quantile = group)

m1<-ggplot(disturbance_spatial_females, aes(x = snow_spatial, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer ~ disturbance:mean snow")


m2<-ggplot(polygon_spatial_females, aes(x = snow_spatial, y = predicted, colour = polygonal_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = polygonal_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer ~ polygonal:mean snow")

m3<-ggplot(linear_spatial_females, aes(x = snow_spatial, y = predicted, colour = linear_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = linear_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer ~ linear:mean snow")

#Snow anomaly

disturbance_anomaly_females<-ggeffects::ggpredict(
             disturbance_females,
  terms = c("snow_anomaly [all]",
            paste0("disturbance_index [", round(disturbance_QL, 2), ",", round(disturbance_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         disturbance_quantile = group)


polygon_anomaly_females<-ggeffects::ggpredict(
             polygonal_females,
  terms = c("snow_anomaly [all]",
            paste0("polygonal_index [", round(polygon_QL, 2), ",", round(polygon_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         polygonal_quantile = group)

linear_anomaly_females<-ggeffects::ggpredict(
             linear_females,
  terms = c("snow_anomaly [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         linear_quantile = group)


m4<-ggplot(disturbance_anomaly_females, aes(x = snow_anomaly, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("femAle deer ~ disturbance:snow anomaly")

m5<-ggplot(polygon_anomaly_females, aes(x = snow_anomaly, y = predicted, colour = polygonal_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = polygonal_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer ~ polygonal:snow anomaly")

m6<-ggplot(linear_anomaly_females, aes(x = snow_anomaly, y = predicted, colour = linear_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = linear_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer ~ linear:snow anomaly")



#Put them all together
ggpubr::ggarrange(m1, m2, m3, m4, m5, m6, nrow = 2, ncol = 3)


```


Young Deer?
```{r}
#All disturbance
disturbance_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_young)

#Polygonal -THIS IS THE TOP MODEL FOR ALL DEER
polygonal_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(polygonal_young)


#Linear
linear_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Predator index
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(linear_young)



#Model selection and cleanup into a nice dataframe!
model_selection_young<- as.data.frame(model.sel(disturbance_young, linear_young, polygonal_young,
              rank = "AICc"))

model_selection_young<-model_selection_young%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_young)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_young)<-NULL


#Model predictions for the polygonal and the linear
polygon_QL<-quantile(wtd_dets$polygonal_index, 0.10, na.rm = TRUE)
polygon_QH<-quantile(wtd_dets$polygonal_index, 0.90, na.rm = TRUE)

linear_QL<-quantile(wtd_dets$linear_index, 0.10, na.rm = TRUE)
linear_QH<-quantile(wtd_dets$linear_index, 0.90, na.rm = TRUE)

disturbance_QL<-quantile(wtd_dets$disturbance_index, 0.10, na.rm = TRUE)
disturbance_QH<-quantile(wtd_dets$disturbance_index, 0.90, na.rm = TRUE)

#Predictions for each disturbance type and snow interaction
disturbance_spatial_young<-ggeffects::ggpredict(
             disturbance_young,
  terms = c("mean_snow_spatial [all]",
            paste0("disturbance_index [", round(disturbance_QL, 2), ",", round(disturbance_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         disturbance_quantile = group)


polygon_spatial_young<-ggeffects::ggpredict(
             polygonal_young,
  terms = c("mean_snow_spatial [all]",
            paste0("polygonal_index [", round(polygon_QL, 2), ",", round(polygon_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         polygonal_quantile = group)

linear_spatial_young<-ggeffects::ggpredict(
             linear_young,
  terms = c("mean_snow_spatial [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_spatial= x,
         linear_quantile = group)

m1<-ggplot(disturbance_spatial_young, aes(x = snow_spatial, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer with young ~ disturbance:mean snow")


m2<-ggplot(polygon_spatial_young, aes(x = snow_spatial, y = predicted, colour = polygonal_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = polygonal_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer with young  ~ polygonal:mean snow")

m3<-ggplot(linear_spatial_young, aes(x = snow_spatial, y = predicted, colour = linear_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = linear_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer with young  ~ linear:mean snow")

#Snow anomaly

disturbance_anomaly_young<-ggeffects::ggpredict(
             disturbance_young,
  terms = c("snow_anomaly [all]",
            paste0("disturbance_index [", round(disturbance_QL, 2), ",", round(disturbance_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         disturbance_quantile = group)


polygon_anomaly_young<-ggeffects::ggpredict(
             polygonal_young,
  terms = c("snow_anomaly [all]",
            paste0("polygonal_index [", round(polygon_QL, 2), ",", round(polygon_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         polygonal_quantile = group)

linear_anomaly_young<-ggeffects::ggpredict(
             linear_young,
  terms = c("snow_anomaly [all]",
            paste0("linear_index [", round(linear_QL, 2), ",", round(linear_QH, 2), "]")),
  type = "fixed")%>%
  rename(snow_anomaly= x,
         linear_quantile = group)


m4<-ggplot(disturbance_anomaly_young, aes(x = snow_anomaly, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("femAle deer with young  ~ disturbance:snow anomaly")

m5<-ggplot(polygon_anomaly_young, aes(x = snow_anomaly, y = predicted, colour = polygonal_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = polygonal_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer with young  ~ polygonal:snow anomaly")

m6<-ggplot(linear_anomaly_young, aes(x = snow_anomaly, y = predicted, colour = linear_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = linear_quantile),
              alpha = 0.15, colour = NA)+
  ggtitle("female deer with young  ~ linear:snow anomaly")



#Put them all together
ggpubr::ggarrange(m1, m2, m3, m4, m5, m6, nrow = 2, ncol = 3)


```






















#################################################################################################

Initial model selection for oil and gas disturbance variables for all deer
```{r}

#Lets build a single variable model for each of the oil and gas features
disturbance_index<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_index)


#Consider going through the roads, this doesn't seem right
roads<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   roads+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(roads)

#Similar with all wells
wells<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   wells+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(wells)

well_gas<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_gas+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_gas)

well_oil<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_oil+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_oil)

well_bitumen<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_bitumen+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_bitumen)

inactive_well<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   inactive_well+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(inactive_well)

seismic<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   seismic+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(seismic)

seismic_3D<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   seismic_3D+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(seismic_3D)


pipeline<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   pipeline+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(pipeline)


#Model selection and cleanup into a nice dataframe!
model.sel(disturbance_index, roads, wells, well_gas, well_oil, well_bitumen, inactive_well, pipeline, seismic, seismic_3D,
              rank = "AICc")

model_selection_disturbance<- as.data.frame(model.sel(disturbance_index, roads, wells, well_gas, well_oil, well_bitumen, inactive_well, pipeline, seismic, seismic_3D,
              rank = "AICc"))

model_selection_disturbance<-model_selection_disturbance%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_disturbance)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_disturbance)<-NULL



#Alright well a bit unsurprising but the disturbance index came out as most informative...
```

adult males
```{r}

#Lets build a single variable model for each of the oil and gas features
disturbance_index<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_index)


#Consider going through the roads, this doesn't seem right
roads<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   roads+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(roads)

#Similar with all wells
wells<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   wells+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(wells)

well_gas<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   well_gas+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_gas)

well_oil<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   well_oil+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_oil)

well_bitumen<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   well_bitumen+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_bitumen)

inactive_well<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   inactive_well+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(inactive_well)

seismic<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   seismic+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(seismic)

seismic_3D<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   seismic_3D+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(seismic_3D)


pipeline<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   pipeline+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(pipeline)


#Model selection and cleanup into a nice dataframe!
model.sel(disturbance_index, roads, wells, well_gas, well_oil, well_bitumen, inactive_well, pipeline, seismic, seismic_3D,
              rank = "AICc")

model_selection_disturbance<- as.data.frame(model.sel(disturbance_index, roads, wells, well_gas, well_oil, well_bitumen, inactive_well, pipeline, seismic, seismic_3D,
              rank = "AICc"))

model_selection_disturbance_male<-model_selection_disturbance%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_disturbance)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_disturbance_male)<-NULL



#Alright well a bit unsurprising but the disturbance index came out as most informative...
```


adult females
```{r}
#Lets build a single variable model for each of the oil and gas features
disturbance_index<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_index)


#Consider going through the roads, this doesn't seem right
roads<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   roads+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(roads)

#Similar with all wells
wells<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   wells+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(wells)

well_gas<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   well_gas+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_gas)

well_oil<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   well_oil+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_oil)

well_bitumen<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   well_bitumen+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_bitumen)

inactive_well<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   inactive_well+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(inactive_well)

seismic<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   seismic+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(seismic)

seismic_3D<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   seismic_3D+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(seismic_3D)


pipeline<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   pipeline+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(pipeline)


#Model selection and cleanup into a nice dataframe!
model_selection_disturbance_females<- as.data.frame(model.sel(disturbance_index, roads, wells, well_gas, well_oil, well_bitumen, inactive_well, pipeline, seismic, seismic_3D,
              rank = "AICc"))

model_selection_disturbance_females<-model_selection_disturbance_females%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_disturbance_females)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_disturbance_females)<-NULL



#Alright well a bit unsurprising but the disturbance index came out as most informative...


```

young!
```{r}
#Lets build a single variable model for each of the oil and gas features
disturbance_index<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(disturbance_index)


#Consider going through the roads, this doesn't seem right
roads<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   roads+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(roads)

#Similar with all wells
wells<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   wells+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(wells)

well_gas<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   well_gas+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_gas)

well_oil<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   well_oil+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_oil)

well_bitumen<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   well_bitumen+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(well_bitumen)

inactive_well<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   inactive_well+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(inactive_well)

seismic<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   seismic+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(seismic)

seismic_3D<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   seismic_3D+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(seismic_3D)


pipeline<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   pipeline+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(pipeline)


#Model selection and cleanup into a nice dataframe!
model_selection_disturbance_young<- as.data.frame(model.sel(disturbance_index, roads, wells, well_gas, well_oil, well_bitumen, inactive_well, pipeline, seismic, seismic_3D,
              rank = "AICc"))

model_selection_disturbance_young<-model_selection_disturbance_young%>%
  # Add model names as a column
  mutate(model = rownames(model_selection_disturbance_young)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

rownames(model_selection_disturbance_young)<-NULL



#Alright well a bit unsurprising but the disturbance index came out as most informative...



```


Lets build the more complicated models now for each
```{r}

#Adult full interaction
m1_adult<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_adult)

#Female full interaction
m1_male<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_male)



#Female full interaction
m1_female<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks - n_adult)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_female)


#Young full interaction
m1_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_young)


#Young full interaction
m1_young_gas<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   well_gas+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_snow_spatial*well_gas+
                   snow_anomaly*well_gas+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_young_gas)


```


With predator index
```{r}

#Adult full interaction
m1_adult<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_adult)

#Female full interaction
m1_male<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_male)



#Female full interaction
m1_female<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_female)


#Young full interaction
m1_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_young)


#Young full interaction
m1_young_gas<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   well_gas+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Wolf index
                   mean_snow_spatial*well_gas+
                   snow_anomaly*well_gas+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_young_gas)


```



With predator index + updated disturbance
```{r}


summary(model_list$all_deer_seismic_3D) #negative effect...
summary(model_list$male_roads)          #POSITIVE!
summary(model_list$female_seismic_3D)   #negative...
summary(model_list$young_seismic_3D)    #negative...

#Adult full interaction
m1_all<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   disturbance_index + #Wolf index
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_all)

#Male full interaction
m1_male<-glmmTMB::glmmTMB(cbind(n_male, (n_weeks - n_male)) ~
                   polygonal_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Wolf index
                   mean_snow_spatial*polygonal_index+
                   snow_anomaly*polygonal_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_male)



#Female full interaction
m1_female<-glmmTMB::glmmTMB(cbind(n_female, (n_weeks - n_female)) ~
                   linear_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Wolf index
                   mean_snow_spatial*linear_index+
                   snow_anomaly*linear_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_female)


#Young full interaction
m1_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_young)


#Young full interaction
m1_young_inactive<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks - n_young)) ~
                   inactive_well+
                   PC1 + PC2 + #Habitat covariates
                   predator_index + #Wolf index
                   mean_snow_spatial*inactive_well+
                   snow_anomaly*inactive_well+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1_young_inactive)


```

I want to see model predictions for the female and young
```{r}

#Quantiles for 3d seismic lines
lowQ  <- quantile(wtd_dets$inactive_well, 0.10, na.rm = TRUE)
highQ <- quantile(wtd_dets$inactive_well, 0.90, na.rm = TRUE)
hist(wtd_dets$seismic_3D)

pred_young<- ggpredict(
  m1_young_inactive,
  terms = c("snow_anomaly [all]",
            paste0("inactive_well [", round(lowQ, 2), ",", round(highQ, 2), "]")),
  type = "fe"
)%>%
  rename(snow_anomaly= x,
         disturbance_quantile = group)


ggplot(pred_young, aes(x = snow_anomaly, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA) +
  labs(
    x = "Scaled Mean Long term snow load",
    y = "Predicted Probability of Deer Presence",
    colour = "Disturbance Index level",
    fill = "Disturbance level",
    title = "All Deer"
  ) +
  scale_colour_manual(values = c("#0072B2", "#D55E00"),
                      labels = c("Low disturbance", "High disturbance")) +
 scale_fill_manual(values = c("#0072B2", "#D55E00"), guide = "none")+
  #                  labels = c("Low disturbance", "High disturbance")) +
  theme_classic(base_size = 14)

```

Model predictions for our models for each group
```{r}
#Adults
summary(m1_young)

#Quantiles for disturbance
lowQ  <- quantile(wtd_dets$disturbance_index, 0.10, na.rm = TRUE)
highQ <- quantile(wtd_dets$disturbance_index, 0.90, na.rm = TRUE)

pred_female<- ggpredict(
  m1_female,
  terms = c("mean_snow_spatial [all]",
            paste0("disturbance_index [", round(lowQ, 2), ",", round(highQ, 2), "]")),
  type = "fe"
)%>%
  rename(mean_snow_spatial = x,
         disturbance_quantile = group)


ggplot(pred_adult, aes(x = mean_snow_spatial, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA) +
  labs(
    x = "Scaled Mean Long term snow load",
    y = "Predicted Probability of Deer Presence",
    colour = "Disturbance Index level",
    fill = "Disturbance level",
    title = "All Deer"
  ) +
  scale_colour_manual(values = c("#0072B2", "#D55E00"),
                      labels = c("Low disturbance", "High disturbance")) +
 scale_fill_manual(values = c("#0072B2", "#D55E00"), guide = "none")+
  #                  labels = c("Low disturbance", "High disturbance")) +
  theme_classic(base_size = 14)


##############################################################
lowQ  <- quantile(wtd_dets$disturbance_index, 0.10, na.rm = TRUE)
highQ <- quantile(wtd_dets$disturbance_index, 0.90, na.rm = TRUE)

pred_female<- ggpredict(
  m1_female,
  terms = c("mean_snow_spatial [all]",
            paste0("disturbance_index [", round(lowQ, 2), ",", round(highQ, 2), "]")),
  type = "fe"
)%>%
  rename(mean_snow_spatial = x,
         disturbance_quantile = group)


ggplot(pred_female, aes(x = mean_snow_spatial, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA) +
  labs(
    x = "Scaled Mean Long term snow load",
    y = "Predicted Probability of Deer Presence",
    colour = "Disturbance Index level",
    fill = "Disturbance level",
    title = "Lone Adult Females"
  ) +
  scale_colour_manual(values = c("#0072B2", "#D55E00"),
                      labels = c("Low disturbance", "High disturbance")) +
 scale_fill_manual(values = c("#0072B2", "#D55E00"), guide = "none")+
  #                  labels = c("Low disturbance", "High disturbance")) +
  theme_classic(base_size = 14)

#################################################
##############################################################
lowQ  <- quantile(wtd_dets$disturbance_index, 0.10, na.rm = TRUE)
highQ <- quantile(wtd_dets$disturbance_index, 0.90, na.rm = TRUE)

pred_young<- ggpredict(
  m1_young,
  terms = c("mean_snow_spatial [all]",
            paste0("disturbance_index [", round(lowQ, 2), ",", round(highQ, 2), "]")),
  type = "fe"
)%>%
  rename(mean_snow_spatial = x,
         disturbance_quantile = group)


ggplot(pred_young, aes(x = mean_snow_spatial, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA) +
  labs(
    x = "Scaled Mean Long term snow load",
    y = "Predicted Probability of Deer Presence",
    colour = "Disturbance Index level",
    fill = "Disturbance level",
    title = "Lone Adult Females"
  ) +
  scale_colour_manual(values = c("#0072B2", "#D55E00"),
                      labels = c("Low disturbance", "High disturbance")) +
 scale_fill_manual(values = c("#0072B2", "#D55E00"), guide = "none")+
  #                  labels = c("Low disturbance", "High disturbance")) +
  theme_classic(base_size = 14)

#################################################
#Quantiles dont work here
lowQ  <- quantile(wtd_dets$well_gas, 0.10, na.rm = TRUE)
highQ <- quantile(wtd_dets$well_gas, 0.90, na.rm = TRUE)

#not enough variation in well_gas
#Need to remove outliers
hist(wtd_dets$well_gas)

#But for now lets use min and max
lowQ<-min(wtd_dets$well_gas)
highQ<-max(wtd_dets$well_gas)

pred_young<- ggpredict(
  m1_young_gas,
  terms = c("snow_anomaly [all]",
            paste0("well_gas [", round(lowQ, 2), ",", round(highQ, 2), "]")),
  type = "fe"
)%>%
  rename(snow_anomaly = x,
         well_gas = group)


ggplot(pred_young, aes(x = snow_anomaly, y = predicted, colour = well_gas)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = well_gas),
              alpha = 0.15, colour = NA) +
  labs(
    x = "Scaled Mean Snow Anomaly",
    y = "Predicted Probability of Deer Presence",
    colour = "Gas Wells",
    fill = "Gas Wells",
    title = "Female with Young"
  ) +
  scale_colour_manual(values = c("#0072B2", "#D55E00"),
                      labels = c("Low", "High")) +
 scale_fill_manual(values = c("#0072B2", "#D55E00"), guide = "none")+
  #                  labels = c("Low disturbance", "High disturbance")) +
  theme_classic(base_size = 14)

```




Next, lets build more complicated models and see what we can come up wiht
```{r}

#Base model 
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1)

#Now lets add in the snow stuff
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_snow_spatial+
                   snow_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1)

#And finally, an interaction...PLEASSSSEEEE
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1)


#Its not clear to me what the temporal dimension gets me
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_snow_spatial*disturbance_index+
                   snow_anomaly*disturbance_index+
                     mean_snow_temporal*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1)


#Repeat the whole thing with temperature?

#Now lets add in the snow stuff
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_min_temp_spatial+
                   temp_anomaly+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1)

#And finally, an interaction...PLEASSSSEEEE
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index+
                   PC1 + PC2 + #Habitat covariates
                   wolf_index + #Wolf index
                   mean_min_temp_spatial*disturbance_index+
                   temp_anomaly*disturbance_index+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1)


```



Lets do a little exercise here, I want to know how these different disturbance metrics are performing
```{r}
#How do I keep losing lat/long??
wtd_dets<-wtd_dets%>%
  left_join(camop%>%
              dplyr::select(site, lat, long),
            by = "site")


#Scale everyone
wtd_dets<-wtd_dets%>%
  mutate(across(c(mean_snow, mean_snow_spatial, snow_anomaly, mean_min_temp, mean_min_temp_spatial, temp_anomaly, disturbance_index, broadleaf, lat, roads, well_gas, well_oil, well_bitumen, inactive_well, seismic, seismic_3D, harvest, pipeline),
              .fns = ~ as.numeric(scale(.x))))  


#Model of the PCs
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   PC1 + PC2+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(m1)


#How does a model behave with broadleaf vs these components
broadleaf_model<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   broadleaf+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(broadleaf_model)




pc_model<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   PC1 + PC2+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(pc_model)

model.sel(broadleaf_model, pc_model, rank = "AICc")


#Global model of disturbance stuff
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index + roads + well_gas + well_oil + well_bitumen + inactive_well + seismic + seismic_3D + harvest + pipeline+ 
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(m1)


#Quick model seleciton exercise?
disturbance_index<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance_index +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(disturbance_index)
roads<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   roads +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(roads)
well_gas<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_gas +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(well_gas)

well_bitumen<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_bitumen +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(well_bitumen)

well_oil<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_oil +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(well_oil)
inactive_well<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   inactive_well + 
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(inactive_well)
seismic<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                  seismic +      (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(seismic)
seismic_3D<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   seismic_3D +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(seismic_3D)
harvest<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   harvest +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(harvest)

pipeline<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   pipeline +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(pipeline)

ms<-model.sel(disturbance_index, roads, well_gas, well_oil, well_bitumen, inactive_well, harvest, pipeline, seismic, seismic_3D,
              rank = "AICc")

ms<- as.data.frame(ms) %>%
  # Add model names as a column
  mutate(model = rownames(ms)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

View(ms)

#Ok well_gas has the most informative and its positive, lets build from there

well_gas<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_gas +
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)

summary(well_gas)


#Snow and snow anomaly?
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_gas + PC1 + PC2 + mean_snow_spatial + snow_anomaly+
                     well_gas*mean_snow_spatial+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)


summary(m1)


m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_gas + PC1 + PC2 + mean_min_temp_spatial + temp_anomaly+
                     well_gas*mean_min_temp_spatial+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)


summary(m1)

#Can we plot this?




m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   well_gas + PC1 + PC2 + mean_min_temp_spatial + temp_anomaly+
                     well_gas*mean_min_temp_spatial+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)


summary(m1)



m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   pipeline + PC1 + PC2 + mean_min_temp_spatial + temp_anomaly+
                     pipeline*mean_min_temp_spatial+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)


summary(m1)



##############################################
#3D seismic has the best predictive effect...but its negative. Strange
#Is there anything that can be done to moderate this?
m1<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   mean_snow_spatial*well_gas+ lat + broadleaf+
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(m1)


lowQ  <- quantile(wtd_dets$well_gas, 0.10, na.rm = TRUE)
highQ <- quantile(wtd_dets$well_gas, 0.90, na.rm = TRUE)


lowQ  <- min(wtd_dets$pipeline)
highQ <- max(wtd_dets$pipeline)

pred_spatial <- ggpredict(
  m1,
  terms = c("mean_min_temp_spatial [all]",
            "pipeline [lowQ, highQ]"),
  type = "fe"
)%>%
  rename(mean_min_temp_spatial= x,
         disturbance_quantile = group)


ggplot(pred_spatial, aes(x = mean_min_temp_spatial, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA) +
  labs(x = "Scaled average long term mean min temp @ site", 
       y = "Probability of deer weekly presence"
  ) +
  scale_colour_manual(values = c("#0072B2", "#D55E00"),
                      labels = c("Low gas wells", "High gas wells")) +
  scale_fill_manual(values = c("#0072B2", "#D55E00"),
                    labels = c("Low gas wells", "High gas wells")) +
  theme_classic(base_size = 14)



```



```{r}
weather_variables<-c("mean_min_temp", "mean_min_temp_spatial", "temp_anomaly",
                     "mean_snow", "mean_snow_spatial", "snow_anomaly")

disturbance_variables<-c("disturbance_index", "roads", "active_well", "inactive_well", "seismic", "seismic_3D",
                         "harvest", "pipeline")


interaction_combos <- expand.grid(weather = weather_variables,
                                  disturbance = disturbance_variables,
                                  stringsAsFactors = FALSE)

# Initialize an empty list to store results
results_list <- list()

# Loop through each combination
for (i in seq_len(nrow(interaction_combos))) {
  w <- interaction_combos$weather[i]
  d <- interaction_combos$disturbance[i]
  
  # Build formula dynamically
  formula_str <- paste0("cbind(deer_present, (n_weeks - deer_present)) ~ ",
                        w, " * ", d, " + (1 | array)")
  formula_obj <- as.formula(formula_str)
  
  # Try fitting the model, skip if it fails
  model <- try(glmmTMB::glmmTMB(formula_obj, family = binomial, data = wtd_dets), silent = TRUE)
  if (inherits(model, "try-error")) next
  
  # Tidy the model output
  tidy_mod <- broom.mixed::tidy(model, effects = "fixed")
  
  # Extract interaction term
  interaction_term <- paste0(w, ":", d)
  interaction_row <- tidy_mod %>% filter(term == interaction_term)
  
  if (nrow(interaction_row) > 0) {
    results_list[[length(results_list) + 1]] <- data.frame(
      weather = w,
      disturbance = d,
      estimate = interaction_row$estimate,
      std_error = interaction_row$std.error,
      z_value = interaction_row$statistic,
      p_value = interaction_row$p.value
    )
  }
}

# Combine results
interaction_results <- bind_rows(results_list)

# Arrange for easy viewing
interaction_results <- interaction_results %>%
  arrange(p_value)

interaction_results


#Testing models
A<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   
                   	
mean_min_temp*active_well + broadleaf
                   
                   +(1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(A)



A<-wtd_dets%>%
  group_by(site)%>%
  summarise(n())

```


Lets do some quick model prep and modeling before lunch break
```{r}



#How do I keep losing lat/long??
wtd_dets<-wtd_dets%>%
  left_join(camop%>%
              dplyr::select(site, lat, long),
            by = "site")

#Lets scale the remaining predictors
wtd_dets<-wtd_dets%>%
  mutate(across(c(mean_snow, mean_snow_spatial, snow_anomaly, mean_min_temp, mean_min_temp_spatial, temp_anomaly, disturbance, broadleaf, lat),
              .fns = ~ as.numeric(scale(.x))))  
         

#I forgot to scale things

#I am sure things are correlated af but I dont care right now
# Select your variables
vars <- wtd_dets %>%
  dplyr::select(mean_snow, mean_snow_spatial, snow_anomaly, mean_min_temp, mean_min_temp_spatial, temp_anomaly, disturbance, broadleaf)

# Compute correlation matrix (use complete.obs to ignore NAs)
cor_matrix <- cor(vars, use = "complete.obs")

# Print the correlation matrix
print(round(cor_matrix, 2))

# Visualize it
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, addCoef.col = "black")
######################################################################################
#Who cares lets build - models of varying winter components
mean_snow<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance +scale(lat) + broadleaf +
                    mean_snow + 
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(mean_snow)


mean_snow_spatial<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance +scale(lat) + broadleaf +
                    mean_snow_spatial + 
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(mean_snow_spatial)


snow_anomaly<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance +scale(lat) + broadleaf +
                    snow_anomaly + 
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(snow_anomaly)



mean_min_temp<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance +scale(lat) + broadleaf +
                    mean_min_temp + 
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(mean_min_temp)


mean_min_temp_spatial<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance +scale(lat) + broadleaf +
                    mean_min_temp_spatial + 
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(mean_min_temp_spatial)


temp_anomaly<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance +scale(lat) + broadleaf +
                    temp_anomaly + 
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(temp_anomaly)


mean_min_temp_spatial_int<-glmmTMB::glmmTMB(cbind(deer_present, (n_weeks - deer_present)) ~
                   disturbance +scale(lat) + broadleaf +
                    mean_min_temp_spatial*disturbance + 
                   (1|array),
                 family = "binomial",
                 data = wtd_dets)
summary(mean_min_temp_spatial_int)


library(MuMIn)
ms<-model.sel(mean_snow, mean_snow_spatial, snow_anomaly, mean_min_temp, mean_min_temp_spatial, temp_anomaly,
          rank = "AICc")

ms<- as.data.frame(ms) %>%
  # Add model names as a column
  mutate(model = rownames(ms)) %>%
  # Keep only the columns you care about
  dplyr::select(model, df, logLik, AICc, delta, weight) %>%
  # Optional: arrange by AICc
  arrange(AICc)%>%
  mutate(weight = round(weight, digits = 2))

View(ms)


###########################
lowQ  <- quantile(wtd_dets$disturbance, 0.10, na.rm = TRUE)
highQ <- quantile(wtd_dets$disturbance, 0.90, na.rm = TRUE)

pred_spatial <- ggpredict(
  mean_min_temp_spatial,
  terms = c("mean_min_temp_spatial [all]",
            paste0("disturbance [", round(lowQ, 2), ",", round(highQ, 2), "]")),
  type = "fe"
)%>%
  rename(mean_min_temp = x,
         disturbance_quantile = group)


ggplot(pred_spatial, aes(x = mean_min_temp, y = predicted, colour = disturbance_quantile)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = disturbance_quantile),
              alpha = 0.15, colour = NA) +
  labs(
    x = "Scaled Mean Minimum Winter Temperature (Spatial)",
    y = "Predicted Probability of Deer Presence",
    colour = "Disturbance level",
    fill = "Disturbance level",
    title = "Interaction between Spatial Mean Winter Temperature and Disturbance"
  ) +
  scale_colour_manual(values = c("#0072B2", "#D55E00"),
                      labels = c("Low disturbance", "High disturbance")) +
  scale_fill_manual(values = c("#0072B2", "#D55E00"),
                    labels = c("Low disturbance", "High disturbance")) +
  theme_classic(base_size = 14)

```




Preparing the data for preliminary models
```{r}
#Lets scale the variables and then explore correlations
temp<-temp%>%
  mutate(across(
    .cols = c(broadleaf, roads, wells, conventional_seismic, pipeline, mean_min_temp, mean_swe, mean_precip, snow_dif, temp_dif, precip_dif, snow_10_mean),
    .fns = ~ as.numeric(scale(.x)),
    .names = "scale_{.col}"
  ))

#Check for correlations
# Select your variables
vars <- temp %>%
  dplyr::select(scale_broadleaf, scale_roads, scale_wells, scale_conventional_seismic, scale_pipeline, scale_mean_min_temp, scale_mean_swe, scale_mean_precip, snow_dif, temp_dif, precip_dif, snow_10_mean)

# Compute correlation matrix (use complete.obs to ignore NAs)
cor_matrix <- cor(vars, use = "complete.obs")

# Print the correlation matrix
print(round(cor_matrix, 2))

# Visualize it
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, addCoef.col = "black")

#Alright precipitation is a bad guy. lets remove it
temp<-temp%>%
  dplyr::select(-c(scale_mean_precip, precip_df)


```


Its all come to down to this. Lets run some models baby
```{r}

#Variance estimates for year were super low, so not including those as a random effect

#global model for adults and young to get started!
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_mean_min_temp + scale_mean_swe +
                   (1|array) + (1|year),
                 family = "binomial",
                 data = temp)
summary(m1)


m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_mean_min_temp + scale_mean_swe +
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)

####################
#interaction terms?
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_mean_min_temp + scale_mean_swe + scale_roads*scale_mean_min_temp +
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)


m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_mean_min_temp + scale_mean_swe + scale_pipeline*scale_mean_min_temp+
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)

###################################
#What about a nice straightforward interaction?
m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_wells*scale_mean_swe+
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)


#What if we account for that north south gradient?
#Adding the coordinates back in
temp<-temp%>%
  left_join(camop%>%
  dplyr::select(site, lat, long),
  by = c("site"))

#Ok yea latitude is important for suresies
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf + lat +
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf + lat + scale_mean_swe +
                       scale_pipeline+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


#Models with the difference stuff
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf + scale_snow_10_mean +
                       scale_pipeline+ scale_snow_dif+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


#Models with the difference stuff
m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf + scale_snow_10_mean + scale_snow_dif +
                       scale_pipeline+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf + scale_snow_10_mean +
                       scale_roads+ scale_snow_dif+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


#Models with the difference stuff
m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf + scale_snow_10_mean + scale_snow_dif +
                       scale_pipeline+
                   (1|array),
                 family = "binomial",
                 data = temp)

summary(m1)


#Global
m1<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_snow_10_mean + scale_snow_dif +
                   (1|array) + (1|year),
                 family = "binomial",
                 data = temp)
summary(m1)

#Global
m1<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_snow_10_mean + scale_snow_dif +
                   (1|array) + (1|year),
                 family = "binomial",
                 data = temp)
summary(m1)



#Global with interaction
m1_adult<-glmmTMB::glmmTMB(cbind(n_adult, (n_weeks-n_adult)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_snow_10_mean + scale_snow_dif + scale_snow_10_mean*scale_roads + scale_snow_dif*scale_roads+
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)

#Global with interaction
m1_young<-glmmTMB::glmmTMB(cbind(n_young, (n_weeks-n_young)) ~ scale_broadleaf +
                   scale_roads + scale_wells + scale_conventional_seismic + scale_pipeline + scale_snow_10_mean + scale_snow_dif +scale_snow_10_mean*scale_roads + scale_snow_dif*scale_roads+
                   (1|array),
                 family = "binomial",
                 data = temp)
summary(m1)

# -----------------------------
# 1. Predicted probabilities for adult-only model
# -----------------------------
roads_q <- quantile(temp$scale_roads, probs = c(0.25, 0.75), na.rm = TRUE)
low_roads <- roads_q[1]
high_roads <- roads_q[2]


adult_pred_low <- ggpredict(m1_adult, terms = c("scale_snow_10_mean [all]", paste0("scale_roads [", round(low_roads, 2), "]")))
adult_pred_low$road_level <- "Low Roads"

adult_pred_high <- ggpredict(m1_adult, terms = c("scale_snow_10_mean [all]", paste0("scale_roads [", round(high_roads, 2), "]")))
adult_pred_high$road_level <- "High Roads"

adult_pred <- bind_rows(adult_pred_low, adult_pred_high)
adult_pred$group <- "Adult Female"

# -----------------------------
# 3. Predicted probabilities for adult+young model
# -----------------------------
young_pred_low <- ggpredict(m1_young, terms = c("scale_snow_10_mean [all]", paste0("scale_roads [", round(low_roads, 2), "]")))
young_pred_low$road_level <- "Low Roads"

young_pred_high <- ggpredict(m1_young, terms = c("scale_snow_10_mean [all]", paste0("scale_roads [", round(high_roads, 2), "]")))
young_pred_high$road_level <- "High Roads"

young_pred <- bind_rows(young_pred_low, young_pred_high)
young_pred$group <- "Adult + Young"

pred_df <- bind_rows(adult_pred, young_pred)

ggplot(pred_df, aes(x = x, y = predicted, color = road_level)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = road_level), alpha = 0.2, color = NA) +
  facet_wrap(~group) +
  labs(
    x = "Long-term Snow Load (10-year mean, scaled)",
    y = "Predicted Probability of Detection",
    color = "Road Density",
    fill = "Road Density"
  ) +
  scale_color_manual(values = c("blue", "red")) +
  scale_fill_manual(values = c("blue", "red")) +
  theme_minimal(base_size = 14)
```
