The purpose of this file is to do the setup of the response variable for deer? Lets do this clean. 
But also predator stuff


```{r}
#Clear everything out and start fresh
rm(list=ls())
library(reproducible) #setting up directories for data management
library(readr)        #Needed to read and merge all the csv's in the camera folder
library(dplyr)        #data wranglin'
library(lubridate)    #datetime manipulation
library(tidyr)        #more data manipulation
library(stringr)      #detecting character strings
library(ggplot2)      #data visualization
library(purrr)        #feline-friendly loops 
library(corrplot)     #correlation tests


#This just sets up files for inputs (already created, must contain your files), and outputs
input_directory<-reproducible::checkPath(file.path(getwd(), "1. Species Response Variables/inputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "1. Species Response Variables/outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "1. Species Response Variables/figures"), create = TRUE)


#read in the files we need
dets<-read.csv(file.path(input_directory, "OSM_camera_detections.csv"))
camop<-read.csv(file.path(input_directory, "OSM_camera_deployment.csv"))
covs<-read.csv(file.path(input_directory, "OSM_camera_covariates.csv"))

#And the weather data
weather<-read.csv(file.path(input_directory, "daymet_weather_data.csv"))

#I am reading this in from the desktop since it is too large to push to github
weather_10<-read.csv("C:/Users/16043/Desktop/OSM white tailed deer/daymet_weather_data_10year.csv")


```


First, we need to calculate independent detections of deer and predator species of interest
```{r}

#What predators do I want? I think lets just go with wolves for now. We can fancy things up later!
unique(dets$species)

#First, I want to know when deer were detected, this is going to influence when and where we look for deer
wtd<-dets%>%
  filter(species %in% c("White-tailed deer", "Grey wolf", "Black bear", "Coyote"))%>%
  #Just ensuring its read in as a datetime here!
  mutate(datetime = ymd_hms(datetime))


#arrange the detection by site, species, and time, then group by site
wtd<-wtd%>%
  arrange(array, site, species, datetime)%>%
  group_by(site)%>%
  #Create a lag time that will be used to calculate time differences
   mutate(duration = as.numeric(difftime(datetime,lag(datetime),units = "mins")))

#Set up the framework for assigning unique IDs
wtd$Event.ID<-9999 #This is just an arbitrary event identification code
mins<-2 # THIS IS THE DETECTION EVENT BREAK-POINT YOU CAN CHANGE
seq <- as.numeric(paste0(nrow(wtd),0))
seq <- round(seq,-(nchar(seq)))

#This will go through each row and figure out how far apart in time each one is
#So this creates unique event identifications for subsequent rows IF they are more than 30mins apart
for (i in 2:nrow(wtd)) {
 wtd$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F))
  if(is.na(wtd$duration[i]) | abs(wtd$duration[i]) > mins){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(wtd$duration[nrow(wtd)] < mins| 
   is.na(wtd$duration[nrow(wtd)])){
  wtd$Event.ID[nrow(wtd)] <- wtd$Event.ID[nrow(wtd)-1] 
} else{
  wtd$Event.ID[nrow(wtd)] <- paste0("E",format(seq+1, scientific = F)) 
}

browseURL("https://www.youtube.com/watch?v=wOFVrjL-XBM")

#Ok I am going to do two things here. Calculate the independent detections of deer and other species normally, but I am going to set aside a dataframe that will just be used for deer demographics. Because there are lots of little things that need to be done for this


#Lets make a dataframe for the deer demographics
wtd_demographics<-wtd%>%
  filter(species == "White-tailed deer")%>%
  dplyr::select(array, site, species, datetime, male, female, yoy, Event.ID)
  

#And then a seperate one for just regular ole independent detections
wtd<-wtd%>%
  dplyr::select(array, site, species, datetime, Event.ID)

#And I think at this point, just take the first observation for each event ID
wtd<-wtd%>%
 group_by(Event.ID)%>%
  slice(1)


```


Data exploration of independent events
```{r}
#How many ind detections
wtd%>%
  group_by(species)%>%
  summarise(n_ind_dets = n())

#How many sites?
n_distinct(wtd$site)
#Site occupancy?
wtd%>%
  group_by(species)%>%
  summarise(n_sites = n_distinct(site),
            naive_occupancy = n_sites/n_distinct(dets$site))


#What do detections look like throughout the year?
#Ok so based on this, a good way to look at young deer would be from June to September each year!
wtd%>%
  mutate(month = month(datetime))%>%
  group_by(month, species)%>%
  summarise(n_obs = n())%>%
  ggplot(aes(x = as.factor(month), y = n_obs, flll = species))+
  geom_bar(stat = "identity", aes(fill = species), position = position_dodge())+
  ylab("Independent Detections")+
  ggtitle("Species Detections ~ Month of Year")+
  xlab("Month")


#Lets create a figure of just white-tailed deer detections throughout the year - this will be a supplementary figure for the manuscript


detections_per_month<-wtd %>%
  filter(species == "White-tailed deer") %>%
  mutate(
    month_num = month(datetime),
    #Adding a variable to colour code the months we are using
    #This is also slick, month.abb?? who knew that was built in!?
    month = factor(month_num, levels = 1:12, labels = month.abb),
    highlight = ifelse(month_num %in% 6:9, "summer", "other")   # Jun–Sep = red
  ) %>%
  group_by(month, highlight) %>%
  summarise(n_obs = n()) %>%
  #And now feed it into a ggplot
  ggplot(aes(x = month, y = n_obs, fill = highlight)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("summer" = "red", "other" = "grey70")) +
  labs(
    x = "Month",
    y = "Total Independent Detections"
  ) +
  theme_classic(base_size = 14) +
  theme(legend.position = "none")

#Save that figure!
ggsave(filename = "detections_per_month.jpeg", #name
        plot = detections_per_month,  #plot you want to save
        path = figure_directory, #where its saved
        width =3000, #how wide
        height = 1500,#how tall
         units = "px", #units in pixels
         bg = 'white') #make sure background is white


```

Lets continue to work on our response variable. A couple things need doing. I am going to stretch this into a weekly proportional binomial for each site. Monthly might be too coarse since only using four months. I need to 1) do this properly such that we only count a week as operational if it ran for all 7 days, and then 2) find a way to fill in true zeros
```{r}

#Lets first work on operability of cameras. We will first use the main detection frame as this has all the days of operation for all sites.
#Filter down to june, july, august, and september of each year
df<-dets%>%
  mutate(month = month(datetime),
         date = date(datetime))%>%
  filter(month %in% 6:9)

#Keep only unique dates for each array/site
df<-df%>%
  distinct(array, site, date)

#Identify the year and week for each site, and only keep sites that had full weeks
weekly_operation<-df%>%
  mutate(year = year(date),
         week = week(date))%>%
  #Have to group by year as well, since some sites ran across several years
  group_by(array, year, site, week)%>%
  summarise(n_days = n_distinct(date))%>%
  #Here we remove sites that didn't have full weeks
  filter(n_days == 7)

#How many sites is this? Nice nice nice
n_distinct(weekly_operation$site)
#How many weeks per site? 
weekly_operation%>%
  group_by(year, site)%>%
  summarise(n_weeks = n_distinct(week))%>%
  summarise(mean_weeks = mean(n_weeks),
            sd_weeks = sd(n_weeks),
            min_weeks = min(n_weeks),
            max_weeks = max(n_weeks))

####################################################################################
#Now that we have a list of cameras with operation we can use, I need to identify how many detentions of each group, for each site/year/week combination

#Lets create a temporary frame to work in so we don't have to re-run the independent detection loop again if needed
wtd_dets<-wtd

#First, filter our dataframe to the same months
wtd_dets<-wtd_dets%>%
  mutate(month = month(datetime))%>%
  filter(month %in% 6:9)

#Add the year and week columns we will need
wtd_dets<-wtd_dets%>%
  mutate(year = year(datetime),
         week = week(datetime))

#Ok now I think I can do this. Since each event with a yoy or lone female is indicated by just a 1 (presence), I should be able to sum occurrences for each year/site/week combination right?
#REMEMBER - this will not account for double absences! Sites with zero young and zero adult will not be captured here. Address this next
# wtd_dets<-wtd_dets%>%
#   group_by(array, year, site, week, species)%>%
#   summarise(n_obs = n())%>%
#   pivot_wider(names_from = species, values_from = n_obs)%>%
#   rename("wtd_present" = "White-tailed deer", "wolf_present" = "Grey wolf")%>%
#   #And then take those sums, and turn them back into presence/absence again
#   #Remember just want to know for each week if a group was present!
#   mutate(across(c(wtd_present, wolf_present),
#                 ~ ifelse(is.na(.x) | .x == 0, 0, 1)))

```


```{r}



#Ok, lets work with the predators for now. At this point I have independent detections of predator species. Lets pull them out, figure out how many per site by just summing across all predators
#REMEMBER - this wont have zeros! This will get filled in later. 
predator_index<-wtd_dets%>%
  filter(species %in% c("Black bear", "Grey wolf", "Coyote"))%>%
  #And because I just want total independent detections, I can just take the number of observations at each site/year/array --> the total over all the weeks!
 group_by(array, year, site, species)%>%
   summarise(predator_detections = n())


#I am keeping that above, but lets do some monkeying around for plotting
predator_index <- wtd_dets %>%
  filter(species %in% c("Black bear", "Grey wolf", "Coyote")) %>%
  group_by(array, year, site, species) %>%
  summarise(predator_detections = n(), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from = species,
    values_from = predator_detections,
    values_fill = 0
  ) %>%
  mutate(
    predator_detections = `Black bear` + `Grey wolf` + Coyote
  )

```



```{r}
#Ok, I want to do something slightly different here.First for deer, I think I can sum the occurrences and turn them into presence/absence
 wtd_dets<-wtd_dets%>%
   filter(species == "White-tailed deer")%>%
   group_by(array, year, site, week)%>%
   summarise(n_obs = n())%>%
   #pivot_wider(names_from = species, values_from = n_obs)%>%
   rename("wtd_present" = "n_obs")%>%
   #And then take those sums, and turn them back into presence/absence again
   #Remember just want to know for each week if a group was present!
   mutate(across(c(wtd_present),
                 ~ ifelse(is.na(.x) | .x == 0, 0, 1)))

#And now, we can merge this with the weekly operation file.
wtd_dets<-left_join(weekly_operation, wtd_dets, by = c("array", "year", "site", "week"))
#And similarly add in the predator_index
#wtd_dets<-left_join(wtd_dets, predator_index, by = c("array", "year", "site"))

#And now, lets fill in the zeros for the weeks where we had no detections
wtd_dets<-wtd_dets%>%
  mutate(wtd_present = ifelse(is.na(wtd_present), 0, wtd_present))

#Math check, how many weeks?
wtd_dets%>%
  group_by(year, site)%>%
  summarise(n_weeks = n_distinct(week))%>%
  ungroup()%>%
  summarise(total_weeks = sum(n_weeks))

#And last step, for each array, site, and year --> how many weeks were sampled, and how many of those weeks had a detection of a young or an adult. Again, since young and adult are coded as present or absent, we can sum them up
wtd_dets<-wtd_dets%>%
  group_by(array, year, site)%>%
  summarise(n_weeks = n_distinct(week),   #how many weeks for each site
            deer_present = sum(wtd_present))#how many of those weeks had a deer
           
#And now I believe we can add in the predator detections and calculate an index
wtd_dets<-left_join(wtd_dets, predator_index, by = c("array", "year", "site"))%>%
  mutate(predator_detections = ifelse(is.na(predator_detections), 0, predator_detections))


#Can we visualize this nicely? Yes, but need to manipulate the frame a little bit for eas
ggplot(wtd_dets, aes(x = array, y = deer_present / n_weeks))+
  geom_jitter()+
  ylab("Proportion Summer Weeks Present")+
  xlab("Array")+
  theme_classic()




#I would like to write this file now for future use!
#write.csv(wtd_dets,file.path(output_directory, "weekly_proportional_binomial.csv"),
#          row.names = FALSE)


```

Quick exploration of the predator detections! Need to formalize this later, clean it up
```{r}
wtd_dets<-wtd_dets%>%
  rename(black_bear = `Black bear`,
         wolf = `Grey wolf`,
         coyote = Coyote)





predator_long <- wtd_dets %>%
  pivot_longer(
    cols = c(black_bear, coyote, wolf, predator_detections),
    names_to = "predator_type",
    values_to = "detections"
  ) %>%
  mutate(
    predator_type = recode(
      predator_type,
      black_bear = "Black bear",
      coyote = "Coyote",
      wolf = "Wolf",
      predator_detections = "Total predators"
    )
  )

#re-order that variable for nice plotting
predator_long$predator_type<-factor(predator_long$predator_type,
                                    levels = c("Black bear",
                                               "Wolf",
                                               "Coyote",
                                               "Total predators"))


ggplot(predator_long, aes(x = detections)) +
  geom_histogram(bins = 30) +
  facet_wrap(~ predator_type, scales = "free_x") +
  theme_minimal() +
  labs(
    x = "Independent detections",
    y = "Frequency",
    title = "Distributions of Predator Detections"
  )


#Quick predator PCA

predator_pca<-wtd_dets %>%
  dplyr::mutate(
    across(
      c(black_bear, wolf, coyote),
      ~ tidyr::replace_na(.x, 0)
    )
  )
predator_pca<-predator_pca%>%
  ungroup()%>%
  dplyr::select(black_bear, wolf, coyote)%>%
    mutate(across(c(black_bear, wolf, coyote),
              .fns = ~ as.numeric(scale(.x))))

pred_pca<-prcomp(predator_pca, center = FALSE, scale. = FALSE)

summary(pred_pca)
pred_pca$rotation

#Ok now remove those three predators from the datafarame, no longer care about them this was just exploration
wtd_dets<-wtd_dets%>%
  dplyr::select(-c(black_bear, wolf, coyote))

```



And now seperately with the previously prepared dataframe, lets work on weekly proportional binomial for the males, female deer, and female with young
```{r}
#Reminder on our dataframe
wtd_demographics

#Second, since we don't care about absolute number of deer in events, we are going to turn things into binary present or absent. 
#First turn those NAs into zeros
wtd_demographics <- wtd_demographics %>%
  mutate(across(c(male, female, yoy),
                ~ ifelse(is.na(.x) | .x == 0, 0, 1)))

#Next, for event ID lets identify if a young was present or if a female was present
wtd_demographics<-wtd_demographics %>%
  group_by(Event.ID) %>%
  mutate(
    # Whether any yoy = 1 within the event
    yoy_present = as.integer(any(yoy == 1, na.rm = TRUE)),
    
    # Whether ONLY females were present (female = 1 somewhere, but no yoy = 1 anywhere)
    female_only = as.integer(any(female == 1, na.rm = TRUE) & !any(yoy == 1, na.rm = TRUE))
  ) %>%
  ungroup()

#Finally, lets reduce to one row per event, and filter out events with zeros for both categories (these would represent males and stuff)
wtd_demographics<-wtd_demographics%>%
  group_by(Event.ID)%>%
  filter(row_number()==1)%>%
  filter(!(yoy_present == 0 & female_only == 0 & male == 0))%>%
  ungroup()%>%
  dplyr::select(-c(female, yoy))%>%
  #Quick, yet inefficient, renaming to female, since we are now dealing with adult males and females
  rename(female = female_only)

#Lets get rid of species we dont need that no mo
wtd_demographics<-wtd_demographics%>%
  dplyr::select(-species)





#First, filter our dataframe to the same months
wtd_demographics<-wtd_demographics%>%
  mutate(month = month(datetime))%>%
  filter(month %in% 6:9)

#Add the year and week columns we will need
wtd_demographics<-wtd_demographics%>%
  mutate(year = year(datetime),
         week = week(datetime))

#Ok now I think I can do this. Since each event with a yoy or lone female is indicated by just a 1 (presence), I should be able to sum occurrences for each year/site/week combination right?
#REMEMBER - this will not account for double absences! Sites with zero young and zero adult will not be captured here. Address this next
wtd_demographics<-wtd_demographics%>%
  group_by(array, year, site, week)%>%
  summarise(sum_yoy = sum(yoy_present),
            sum_female = sum(female),
            sum_male = sum(male))%>%
  #And then take those sums, and turn them back into presence/absence again
  #Remember just want to know for each week if a group was present!
  mutate(across(c(sum_yoy, sum_female, sum_male),
                ~ ifelse(is.na(.x) | .x == 0, 0, 1)))%>%
  #rename for clarity
  rename(yoy_present = sum_yoy,
         female_present = sum_female,
         male_present = sum_male)


#And now, we can merge this with the weekly operation file.
wtd_demographics<-left_join(weekly_operation, wtd_demographics, by = c("array", "year", "site", "week"))

#And now, lets fill in the zeros for the weeks where neither a young or adult was detected
wtd_demographics<-wtd_demographics%>%
  mutate(yoy_present = ifelse(is.na(yoy_present), 0, yoy_present),
         female_present = ifelse(is.na(female_present), 0, female_present),
         male_present = ifelse(is.na(male_present), 0, male_present))

#Math check, how many weeks?
wtd_demographics%>%
  group_by(year, site)%>%
  summarise(n_weeks = n_distinct(week))%>%
  ungroup()%>%
  summarise(total_weeks = sum(n_weeks))

#And last step, for each array, site, and year --> how many weeks were sampled, and how many of those weeks had a detection of a young or an adult. Again, since young and adult are coded as present or absent, we can sum them up
wtd_demographics<-wtd_demographics%>%
  group_by(array, year, site)%>%
  summarise(n_weeks = n_distinct(week),   #how many weeks for each site
            n_young = sum(yoy_present),   #how many of those weeks had a young
            n_female = sum(female_present),
            n_male = sum(male_present)) #how many of those weeks had an adult



#Can we visualize this nicely? Yes, but need to manipulate the frame a little bit for eas
wtd_demographics%>%
  pivot_longer(cols = c(n_young, n_female, n_male),
               values_to = "n_present")%>%
ggplot(aes(x = array, y = n_present / n_weeks, color = name))+
  geom_jitter()+
  ylab("Proportion Summer Weeks Present")+
  xlab("Array")+
  scale_color_discrete(
    name = "Age Class",
    labels = c("n_young" = "Fawns", "n_female" = "Females", "n_male" = "Males"))+
  theme_classic()



#And NOW, we should be able to merge with the main file!
wtd_dets<-wtd_dets%>%
  left_join(wtd_demographics, by = c("array", "site", "year", "n_weeks"))


#write that file baby
write.csv(wtd_dets,file.path(output_directory, "weekly_proportional_binomial_with_males.csv"),
         row.names = FALSE)
```

Response Variable Plot
```{r}

wtd_long<-wtd_dets %>%
  #First lets make our new columns to plot the actual proportion
  mutate(
    prop_males   = n_male / n_weeks,
    prop_females = n_female / n_weeks,
    prop_young   = n_young / n_weeks
  ) %>%
  #Clean up, dataframe 2 messy 4 me
  dplyr::select(array, year, site, prop_males, prop_females, prop_young) %>%
  tidyr::pivot_longer(
    cols = c(prop_males, prop_females, prop_young),
    names_to = "group",
    values_to = "proportion"
  )
  
#Build, save, and export the figure
response_variable_figure<-ggplot(wtd_long, aes(x = array, y = proportion, fill = group)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  scale_fill_brewer(palette = "Set2",
                    labels = c("Males", " Lone Females", "Females With\nYoung")) +
  labs(
    x = "Array",
    y = "Proportion of Weeks With Detections",
    fill = "Group"
  ) +
  theme_classic(base_size = 14)


ggsave(filename = "response_variable_figure.jpeg", #name
        plot = response_variable_figure,  #plot you want to save
        path = figure_directory, #where its saved
        width =3000, #how wide
        height = 1500,#how tall
         units = "px", #units in pixels
         bg = 'white') #make sure background is white

```
